---
title: "PRV - 3 cell line: TES dynamics"
author: Balázs Kakuk
output: 
  html_document:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: false
---

```{css, echo=FALSE}
    body .main-container {
      max-width: 90% !important;
      width: 90% !important;
    }
    body {
      max-width: 90% !important;
      margin-left: auto;
      margin-right: auto;
    }
```


```{r include=FALSE}
## hrbrthemes::ipsum:

knitr::opts_chunk$set(fig.retina=2, echo = FALSE, fig.align = 'center', message = F, warning = F)
```

```{r ipsum_setup, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}
library(prettydoc)
library(hrbrthemes, quietly = T)
#library(GenomicFeatures, quietly = T)
library(DESeq2, quietly = T)
library(ggpubr, quietly = T)
#library(misc, quietly = T)
library(dplyr, quietly = T)
#library(tidyverse, quietly = T)
library(ggsci, quietly = T)
#library(Gviz, quietly = T)
library(dplyr, quietly = T)
library(tidyr)
library(scales)
#library(plyr)
library(moanin, quietly = T)
library(grid, quietly = T)
library(gridExtra, quietly = T) 
library(ggpubr, quietly = T)
library(data.table)
library(ggh4x)
library(knitr)
library(formattable)


###
source('_WF.part0.R')

#### Settings ####

## Main
#save.data <- 'PRV.rebase.RData'
EndType <- 'TES'

project_config  <- fread('project_config.txt')
outdir  <- project_config$outdir
res.dir <- outdir; try({ dir.create(res.dir) })
fig.dir <- paste0(res.dir, '/', EndType, '_abund'); # try({ dir.create(fig.dir) })## save plots to this directory

## Miscallenaous
palette <- pal_npg()(10)

writetables <- T



```


# Project info
*Samples: dcDNA Seq*  
*pychopper: no*  
*mapping version: v6*  
*LoRTIA: yes (stranded only)*  

*Filtering samples: C6_2h_3 & PC-12_8h_3*  

*Normalization: host_read.count*  
*Norm. multiplier: 10000*  

## Only those reads were counted, where LoRTIA   
found the *3-prime* adapters!


```{r, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}

#### Import results of all Normalization methods

##
norm.method   <- 'average_genome'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_genome', '/', norm.method, '_', EndType, '_abund.mean.tsv'))
orf.perc.mean[,norm.method := norm.method]
orf.mean.average_genome <- orf.perc.mean

orf.sum.perc <- fread(paste0(fig.dir, '.norm_genome', '/', norm.method, '_', EndType, '_counts.norm.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.average_genome <- orf.sum.perc

##
norm.method   <- 'viral_read.count'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_LoRTIA', '/', norm.method, '_', EndType, '_abund.mean.tsv'))
orf.perc.mean[,norm.method := norm.method]
orf.mean.viral_read.count <- orf.perc.mean

orf.sum.perc  <- fread(paste0(fig.dir, '.norm_LoRTIA', '/', norm.method, '_', EndType, '_counts.norm.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.viral_read.count <- orf.sum.perc

##
norm.method   <- 'total_read.count'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_total', '/', norm.method, '_', EndType, '_abund.mean.tsv'))
orf.perc.mean[,norm.method := norm.method]
orf.mean.total_read.count <- orf.perc.mean

orf.sum.perc <- fread(paste0(fig.dir, '.norm_total', '/', norm.method, '_', EndType, '_counts.norm.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.total_read.count <- orf.sum.perc

##
norm.method   <- 'host_read.count'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_host', '/', norm.method, '_', EndType, '_abund.mean.tsv'))
orf.perc.mean[,norm.method := norm.method]
orf.mean.host_read.count <- orf.perc.mean

orf.sum.perc <- fread(paste0(fig.dir, '.norm_host', '/', norm.method, '_', EndType, '_counts.norm.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.host_read.count <- orf.sum.perc

##
norm.method   <- 'WO'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_Without', '/', norm.method, '_', EndType, '_abund.mean.tsv')) 
orf.perc.mean[,norm.method := norm.method]
orf.mean.raw_read.count <- orf.perc.mean

orf.sum.perc <- fread(paste0(fig.dir, '.norm_Without', '/', norm.method, '_', EndType, '_counts.norm.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.raw_read.count <- orf.sum.perc


###
orf.perc.mean <- rbind(
  orf.mean.average_genome,
  orf.mean.viral_read.count,
  orf.mean.total_read.count,
  orf.mean.host_read.count,
  orf.mean.raw_read.count
)

orf.perc.mean[,Kinetic_class := factor(Kinetic_class, levels = c('IE', 'IE/E', 'E', 'IE/E/L', 'E/L', 'L', 'unknown'))]

###
orf.perc.sum <- rbind(
  orf.sum.average_genome,
  orf.sum.viral_read.count,
  orf.sum.total_read.count,
  orf.sum.host_read.count,
  orf.sum.raw_read.count
)

orf.perc.sum[,Kinetic_class := factor(Kinetic_class, levels = c('IE', 'IE/E', 'E', 'IE/E/L', 'E/L', 'L', 'unknown'))]


### Summarise based on literature kinetic class
kin.class.sum  <- orf.perc.sum[,.(sum_read_count=sum(read_count), mean_read_count=mean(read_count), sd_read_count=sd(read_count)), 
                              by=.(Kinetic_class, hpi, Time, cell_line, seqnames, norm_base, sample, group, norm.method)]


### Summarise based on SIGNIFICANT de-novo kinetic cluster
sig.cluster.sum <- orf.perc.sum[,.(sum_read_count=sum(read_count), mean_read_count=mean(read_count), sd_read_count=sd(read_count)), 
                              by=.(labels, hpi, Time, cell_line, seqnames, norm_base, sample, group, norm.method)]


### Summarise based on all de-novo kinetic cluster
all.cluster.sum <- orf.perc.sum[,.(sum_read_count=sum(read_count), mean_read_count=mean(read_count), sd_read_count=sd(read_count)), 
                              by=.(best_label, hpi, Time, cell_line, seqnames, norm_base, sample, group, norm.method)]



###
norm.method <- 'norm_Combined'
fig.dir <- paste0(fig.dir, '.', norm.method); dir.create(fig.dir)

```


# 2. TES dynamics (matching the 3-prime ends of reads to the *canonic* TES of genes)


## Gene dynamics, according to literature-based kinetic classes {.tabset}
Time-dependent normalized gene counts (based TES clusters)

```{r}
### ORFs in samples calculated from TRS-B transcripts with and without leaders


plotfun <- function(DT, 
                    geom1 = geom_pointrange(aes(ymin = mean-sd, ymax = mean+sd, colour = cell_line)),
                    geom2 = geom_smooth(aes(x=Time, y=mean, color=cell_line)),
                    ncol = 10, title = NULL, #"Normalized ORF counts", 
                    strip.backgr = pal_d3()(10)[2]) {
  
  if (nrow(DT) == 0) { 
    ggp <- NULL } else { # try({ 
      
    # 
    gene_N <- length(unique(DT[,gene]))
    # class_freq <- DT[best_label == 5][,.N,by=class_col]
    
    #colid <- c('gene', class_col)
    #class_freq <- unique(DT[best_label == 5, ..colid])
    
    
    #  min(class_freq)
     # =='adaptive') {}
        
    ggp <- 
      ggplot(DT[
                                #  orf.perc.mean$hpi    != 'dRNA' 
                                #& orf.perc.mean$strand == '+'
                                , ], 
                     aes(Time, mean, color = 'cell_line')
                     #, add = 'loess'
                     ) + 
      geom1 +
      geom2 + 
      scale_color_manual(values = palette) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      # facet_wrap(~ gene_cluster + ID, scales = 'free_y', ncol = ncol) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      ggtitle(title) +
      facet_nested(rows=vars(Kinetic_class, gene), scales = "free_y") ## cols=vars(norm.method)
  
  if (gene_N < ncol) { 
    ggp <- cowplot::plot_grid(ggp, NULL, rel_widths = c(gene_N, ncol - gene_N ), nrow=1)
  }
  
  }   # })
  
  return(ggp)

}

IE.ratio <- 0.45
ncol     <- 10
relh     <- c(1,2,6)
figw     <- 20
figh     <- 120

```


### Mean and SD with linear connection
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}
ggp <- suppressWarnings(cowplot::plot_grid(
  plotfun(orf.perc.mean[norm.method == 'average_genome'],     title = 'Average viral genome coverage', geom2 = geom_line(aes(x=Time, y=mean, color=cell_line)), strip.backgr = pal_d3()(10)[1]) + theme(legend.position = 'none', strip.text.y = element_blank()),
  plotfun(orf.perc.mean[norm.method == 'viral_read.count'] ,  title = 'Viral read count',      geom2 = geom_line(aes(x=Time, y=mean, color=cell_line)), strip.backgr = pal_d3()(10)[2]) +         theme(legend.position = 'none', strip.text.y = element_blank()),
  #plotfun(orf.perc.mean[norm.method == 'total_read.count'] ,  title = 'Total read count',      geom2 = geom_line(aes(x=Time, y=mean, color=cell_line)), strip.backgr = pal_d3()(10)[3]) +         theme(legend.position = 'none', strip.text.y = element_blank()),
  plotfun(orf.perc.mean[norm.method == 'host_read.count'] ,   title = 'Host read count',       geom2 = geom_line(aes(x=Time, y=mean, color=cell_line)), strip.backgr = pal_d3()(10)[4]) +         theme(legend.position = 'none', strip.text.y = element_blank()),
  plotfun(orf.perc.mean[norm.method == 'WO'] ,                title = 'Without normalization', geom2 = geom_line(aes(x=Time, y=mean, color=cell_line)), strip.backgr = pal_d3()(10)[5]) +         theme(legend.position = 'right'),
  nrow = 1, 
  rel_widths = c(1,1,1,1.5)
))

ggp

ggsave(paste0(fig.dir, '/', EndType, '_CombinedNorm.mean_sd_linear.jpg'), ggp, width = figw, height = figh, limitsize = F)

```


### Mean and SD with loess function
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- suppressWarnings(cowplot::plot_grid(
  plotfun(orf.perc.mean[norm.method == 'average_genome'],     title = 'Average viral genome coverage', strip.backgr = pal_d3()(10)[1]) + theme(legend.position = 'none', strip.text.y = element_blank()),
  plotfun(orf.perc.mean[norm.method == 'viral_read.count'] ,  title = 'Viral read count',      strip.backgr = pal_d3()(10)[2]) +       theme(legend.position = 'none', strip.text.y = element_blank()), 
  #plotfun(orf.perc.mean[norm.method == 'total_read.count'] ,  title = 'Total read count',      strip.backgr = pal_d3()(10)[3]) +       theme(legend.position = 'none', strip.text.y = element_blank()), 
  plotfun(orf.perc.mean[norm.method == 'host_read.count'] ,   title = 'Host read count',       strip.backgr = pal_d3()(10)[4]) +       theme(legend.position = 'none', strip.text.y = element_blank()), 
  plotfun(orf.perc.mean[norm.method == 'WO'] ,                title = 'Without normalization', strip.backgr = pal_d3()(10)[5]) +       theme(legend.position = 'right'), 
  nrow = 1, 
  rel_widths = c(1,1,1,1.5)
))

ggp

ggsave(paste0(fig.dir, '/', EndType, '_CombinedNorm.mean_sd_loess.jpg'), ggp, width = figw, height = figh, limitsize = F)

```

## Compare literature and de novo kinetic classes

*NA shows the number of genes where the clustering did not yield an unambiguous result.*
  
```{r, fig.show='hold',  fig.width = 13, fig.height = 13, eval=T}
setDT(orf.perc.mean)
class.dt <- unique(orf.perc.mean[, .(norm.method, gene, Kinetic_class, best_label, labels)])
class.dt <- class.dt[, .N, by=.(norm.method, Kinetic_class, best_label, labels)]

#cluster_labels

ggplot(class.dt,
       aes(x=Kinetic_class, y=N, fill=as.factor(best_label))) +
  geom_col(position = 'dodge') +
  scale_color_manual(values = palette) +
  #scale_y_continuous(labels = scales::percent_format()) +
  theme_ipsum() +
  facet_nested(cols=vars(labels), rows=vars(norm.method))


ggplot(class.dt,
       aes(x=as.factor(best_label), y=N, fill=Kinetic_class)) +
  geom_col(position = 'stack') +
  scale_color_manual(values = palette) +
  #scale_y_continuous(labels = scales::percent_format()) +
  theme_ipsum() +
  facet_nested(cols=vars(labels), rows=vars(norm.method))


```


```{r}
figw <- 15
figh <- 12
ncol <- 10
```



### De Novo - Significant
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- 
      ggplot(sig.cluster.sum,
             aes(Time, mean_read_count, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_point() +
      geom_smooth() + 
      scale_color_manual(values = palette[7:9]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      facet_nested(rows=vars(norm.method), cols=vars(labels), scales = "free_y") ## cols=vars(norm.method)
      # facet_wrap(~ labels, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.DeNovo.Sig.Clusters.', '_points.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)

```


### De Novo - All clusters
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- 
      ggplot(all.cluster.sum,
             aes(Time, mean_read_count, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_point() +
      geom_smooth() + 
      scale_color_manual(values = palette[7:9]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      facet_nested(rows=vars(norm.method), cols=vars(best_label), scales = "free_y") ## cols=vars(norm.method)
      #facet_wrap(~ best_label, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.DeNovo.All.Clusters.', '_points.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)

```


### Literature clusters
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- 
      ggplot(kin.class.sum,
             aes(Time, mean_read_count, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_point() +
      geom_smooth() + 
      scale_color_manual(values = palette[7:9]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      facet_nested(rows=vars(norm.method), cols=vars(Kinetic_class), scales = "free_y") ## cols=vars(norm.method)
      #facet_wrap(~ Kinetic_class, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.Literature.Clusters.', '_points.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)

```


## Literature Clusters and Selected Genes {.tabset}
```{r}
### Summarise based on literature kinetic class
select.genes.class.sum <- 
  orf.perc.sum[!gene %in% c('US1', 'EP0', 'IE180'),
               .(sum_read_count=sum(read_count), mean_read_count=mean(read_count), sd_read_count=sd(read_count)), 
               by=.(Kinetic_class, hpi, Time, cell_line, seqnames, norm_base, sample, group, norm.method)]

select.genes.class.sum <- rbind(
  select.genes.class.sum,
  orf.perc.sum[gene %in% c('US1', 'EP0', 'IE180'),
               .(sum_read_count=sum(read_count), mean_read_count=mean(read_count), sd_read_count=sd(read_count)), 
               by=.(gene, hpi, Time, cell_line, seqnames, norm_base, sample, group, norm.method)],
  fill = T
)


select.genes.class.sum[gene %in% c('US1', 'EP0', 'IE180'), Kinetic_class := gene]


select.genes.class.sum[,mean_cell_line := mean(mean_read_count),
                       by=.(seqnames, Kinetic_class, hpi, Time, cell_line,group,  norm.method)]

select.genes.class.sum[,sd_cell_line   := sd(mean_read_count),
                       by=.(seqnames, Kinetic_class, hpi, Time, cell_line,group,  norm.method)]


select.genes.class.cell.lines <- unique(select.genes.class.sum[, 
            .(mean_cell_line, sd_cell_line, seqnames, Kinetic_class, hpi, Time, cell_line,group,  norm.method)] )

```

### Reps with loess and confidence
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- 
      ggplot(select.genes.class.sum[!norm.method %in% c('total_read.count', 'WO')],
             aes(Time, mean_read_count, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_point() +
      geom_smooth() + 
      scale_color_manual(values = colorvec[12:14]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      facet_nested(rows=vars(norm.method), cols=vars(Kinetic_class), scales = "free_y") ## cols=vars(norm.method)
      #facet_wrap(~ Kinetic_class, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.Literature.Clusters.SELECT', '_points.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)
```

### Means with SD and linear connection
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- 
      ggplot(select.genes.class.cell.lines[!norm.method %in% c('total_read.count', 'WO')],
             aes(Time, mean_cell_line, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_pointrange(aes(Time, ymin=mean_cell_line - sd_cell_line, ymax=mean_cell_line + sd_cell_line, color = cell_line)) +    
      geom_point() +
      geom_line(aes(group = cell_line)) + 
      scale_color_manual(values = colorvec[12:14]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      facet_nested(rows=vars(norm.method), cols=vars(Kinetic_class), scales = "free_y") ## cols=vars(norm.method)
      #facet_wrap(~ Kinetic_class, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.Literature.Clusters.SELECT', '_mean.SD.linear.jpg'), ggp, width = figw, height = figh, limitsize = F)
```



## Gene-pair ratios

These pairs were analyzed: 
ie180/us1, ie180/ep0, us1/ep0, noir1/noir2, noir1/ie180, noir2/ie180

```{r, fig.show='hold',  fig.width = 10, fig.height = 7, eval=F}




ab.US1    <- setnames(orf.perc.sum[gene == 'US1',    .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'US1')

ab.IE180  <- setnames(orf.perc.sum[gene == 'IE180',  .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'IE180')

ab.EP0    <- setnames(orf.perc.sum[gene == 'EP0',    .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'EP0')

ab.noir   <- setnames(orf.perc.sum[gene == 'NOIR',   .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'NOIR')

ab.noir2  <- setnames(orf.perc.sum[gene == 'NOIR-2', .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'NOIR-2')


ab.merged <- merge(ab.US1,    ab.IE180,  by=c(metacols, 'norm.method'))
ab.merged <- merge(ab.merged, ab.EP0,    by=c(metacols, 'norm.method'))
ab.merged <- merge(ab.merged, ab.noir,   by=c(metacols, 'norm.method'))
ab.merged <- merge(ab.merged, ab.noir2,  by=c(metacols, 'norm.method'))

ab.merged[, IE180_US1   :=  IE180   /  US1]
ab.merged[, IE180_EP0   :=  IE180   /  EP0]
ab.merged[, US1_EP0     :=  US1     /  EP0]
#ab.merged[, NOIR1_NOIR2 :=  NOIR    / `NOIR-2`]
#ab.merged[, NOIR_IE180  :=  NOIR    /  IE180]
ab.merged[, NOIR2_IE180 := `NOIR-2` /  IE180]

#ab.merged[, US1_IE180   :=  US1     / IE180     ]
#ab.merged[, EP0_IE180   :=  EP0     / IE180     ]
#ab.merged[, EP0_US1     :=  EP0     / US1       ]
ab.merged[, NOIR2_NOIR1 := `NOIR-2` / NOIR      ]
ab.merged[, IE180_NOIR  :=  IE180   / NOIR      ]
#ab.merged[, IE180_NOIR2 :=  IE180   / `NOIR-2`  ]

gpairs <- c(
'IE180_US1' , 
'IE180_EP0'  , 
'US1_EP0'     ,
'NOIR1_NOIR2' ,
'NOIR_IE180'  ,
'NOIR2_IE180' ,

'US1_IE180'   ,
'EP0_IE180'   ,
'EP0_US1'     ,
'NOIR2_NOIR1' ,
'IE180_NOIR'  ,
'IE180_NOIR2' 
)
gpairs <- gsub('_', '/', gpairs)

ab.merged.gt <- ab.merged[,-c(7:11)]

ab.merged.gt <- melt(ab.merged.gt, variable.name = 'gene_pair', value.name = 'ratio', id.vars = c(1:6))

ab.merged.gt$gene_pair <- gsub('_', '/', ab.merged.gt$gene_pair)
ab.merged.gt[,gene_pair := factor(gene_pair, levels = gpairs)]

ab.merged.gt[,mean := mean(ratio),   by=.(hpi, Time, cell_line, group, norm.method, gene_pair)]
ab.merged.gt[,sd   := sd  (ratio),   by=.(hpi, Time, cell_line, group, norm.method, gene_pair)]
ab.merged.gt[,sum  := sum (ratio),   by=.(hpi, Time, cell_line, group, norm.method, gene_pair)]



plot_data <- ab.merged.gt[norm.method == 'viral_read.count'
                 # norm.method == 'host_read.count' 
                 , ]

ggmean <- ggplot(plot_data) + 
  #geom_point(aes(Time, ratio, color=cell_line)) +
  geom_line(aes(Time, mean, color=cell_line, group=cell_line)) +
  geom_pointrange(aes(Time, mean, color=cell_line, group=cell_line, ymin=mean-sd, ymax=mean+sd)) +
  theme_bw() +
  #facet_nested(rows = vars(norm.method), cols=vars(Time))
  facet_wrap(~gene_pair, scales = 'free', ncol=3)


ggmean
ggsave(paste0(fig.dir, '/gene_ratios_mean.jpg'), width = 10, height = 7)


ggrep <- ggplot(plot_data[# norm.method == 'viral_read.count'
                 # norm.method == 'host_read.count' 
                 , ]) + 
  geom_point(aes(Time, ratio, color=cell_line)) +
  geom_smooth(aes(Time, ratio, color=cell_line, group=cell_line)) +
  #geom_pointrange(aes(Time, mean, color=cell_line, group=cell_line, ymin=mean-sd, ymax=mean+sd)) +
  theme_bw() +
  #facet_nested(rows = vars(norm.method), cols=vars(Time))
  facet_wrap(~gene_pair, scales = 'free')


ggrep
ggsave(paste0(fig.dir, '/gene_ratios_rep.jpg'), width = 10, height = 7)


ggsum <- ggplot(plot_data) + 
  geom_point(aes(Time, sum, color=cell_line)) +
  geom_line(aes(Time, sum, color=cell_line, group=cell_line)) +
  #geom_pointrange(aes(Time, mean, color=cell_line, group=cell_line, ymin=mean-sd, ymax=mean+sd)) +
  theme_bw() +
  #facet_nested(rows = vars(norm.method), cols=vars(Time))
  facet_wrap(~gene_pair, scales = 'free', ncol=3)


ggsum
ggsave(paste0(fig.dir, '/gene_ratios_sum.jpg'), width = 10, height = 7)

```





```{r, eval=F}
## Export data
if (writetables) {
  
  fwrite(orf.perc.sum,  paste0(fig.dir, '/', norm.method, '_', EndType, '_kin.class.sum.tsv'))  
  fwrite(orf.perc.sum,  paste0(fig.dir, '/', norm.method, '_', EndType, '_sig.cluster.sum.tsv')) 
  fwrite(orf.perc.sum,  paste0(fig.dir, '/', norm.method, '_', EndType, '_all.cluster.sum.tsv')) 
  fwrite(orf.perc.sum,  paste0(fig.dir, '/', norm.method, '_', EndType, '_counts.norm.tsv'))
  fwrite(orf.perc.mean, paste0(fig.dir, '/', norm.method, '_', EndType, '_abund.mean.tsv'))
  
}
```


```{r bib, include=FALSE}
# KEEP THIS AT THE END OF THE DOCUMENT TO GENERATE A LOCAL bib FILE FOR PKGS USED
knitr::write_bib(sub("^package:", "", grep("package", search(), value=TRUE)), file='skeleton.bib')
```