---
title: "PRV - 4 cell lines: Transcript (TSS ~ TES) dynamics"
author: Balázs Kakuk
output: 
  html_document:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: false
---

```{css, echo=FALSE}
    body .main-container {
      max-width: 90% !important;
      width: 90% !important;
    }
    body {
      max-width: 90% !important;
      margin-left: auto;
      margin-right: auto;
    }
```


```{r include=FALSE}
## hrbrthemes::ipsum:

knitr::opts_chunk$set(fig.retina=2, echo = FALSE, fig.align = 'center', message = F, warning = F)
```

```{r ipsum_setup, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE, include=FALSE}
library(prettydoc)
library(hrbrthemes, quietly = T)
#library(GenomicFeatures, quietly = T)
library(DESeq2, quietly = T)
library(ggpubr, quietly = T)
#library(misc, quietly = T)
library(dplyr, quietly = T)
#library(tidyverse, quietly = T)
library(ggsci, quietly = T)
#library(Gviz, quietly = T)
library(dplyr, quietly = T)
library(tidyr)
library(scales)
#library(plyr)
library(moanin, quietly = T)
library(grid, quietly = T)
library(gridExtra, quietly = T) 
library(ggpubr, quietly = T)
library(data.table)
library(ggh4x)
library(knitr)
library(formattable)



### Project name and output directory
outdir  <- 'PRV-MDBIO-4cell'; try({ dir.create(outdir) })

#### Load configuration file
config <- readRDS(paste0(outdir, '_config.rds'))

####
source('_WF.part0.R')

#### Settings ####

## Main
#save.data <- 'PRV.rebase.RData'
EndType <- 'TSS.TES'

#project_config  <- fread('project_config.txt')
#outdir  <- project_config$outdir
res.dir <- outdir; try({ dir.create(res.dir) })
fig.dir <- paste0(res.dir, '/', EndType, '_abund'); # try({ dir.create(fig.dir) })## save plots to this directory

## Miscallenaous
palette <- pal_npg()(10)

writetables <- T


## Filter metadata for those samples that had at least 1 viral read
readcounts <- fread(paste0(outdir, '/readcounts.tsv'), header = T)
metafilt   <- metafilt[metafilt$sample %in% readcounts$sample, ]

```


# Project info
Samples: dcDNA Seq  
pychopper: no  
mapping version: v6  
LoRTIA: yes (stranded only output)  

Filtering samples: C6_2h_3 & PC-12_8h_3    

Normalization: Viral read count

## Only those reads were counted, where LoRTIA found *5-prime* adapters and *3-prime polyA* sequences !

```{r, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE, eval=T}
#### Import TSS counts

gene.sample_count.sp <- fread(paste0(res.dir, '/gene.sample.count.sp.tsv'), na.strings = '')


#### !!! Consider adapters!
adapters <- 'both'

source('Gene.counts.R')

#gene.sample_count    <- TR.ref.sum[,.(count=sum(read_count)), by=.(seqnames, strand, gene_cluster, gene, Kinetic_class, sample)]
#gene.sample_count.sp <- dcast(gene.sample_count, seqnames + strand + gene_cluster + gene + Kinetic_class ~ sample)
####

orf.sum <- gene.sample_count.sp

## Filter out multi-copied genes
orf.sum <- orf.sum[!gene %in% c('IE180_2', 'US1_2', 'NOIR2', 'NOIR2-2'),]
orf.sum <- orf.sum[!gene_cluster %in% c('IE180_2', 'US1_2', 'NOIR2', 'NOIR2-2'),]

## Format columns
#orf.sum <- orf.sum[,!c("seqnames", "gene_cluster"), with = FALSE]
#setnames(orf.sum, old=c('Kinetic_class'), new=c('ID'))
orf.sum[,ID := gene]

# Reordering columns to make new_column the first column
setcolorder(orf.sum, c("ID", setdiff(names(orf.sum), "ID")))

## Factorize
orf.sum$gene          <- factor(orf.sum$gene,  levels=ORFs)
orf.sum$ID            <- factor(orf.sum$ID,    levels=ORFs)
orf.sum$Kinetic_class <- factor(orf.sum$Kinetic_class, levels=c('IE', 'E', 'L'))

## Order
orf.sum        <- orf.sum[order(Kinetic_class, gene),]

##
orf.counts     <- as.data.frame(orf.sum)

orf.sum        <- merge(gather(orf.sum, sample, read_count, -c(1:6)), metafilt[,metacols], by='sample')
setDT(orf.sum)



#### Normalization method
norm.method <- 'viral_read.count'
multip      <- 10000

 if (norm.method == 'average_genome') {
  
  ## Average genome norm method
  norm_cov_summary <- fread('norm.cov.summary.tsv')
  normBase <- norm_cov_summary[,.(sample, average_coverage)]
  colnames(normBase)[2] <- 'norm_base'
  
  fig.dir <- paste0(fig.dir, '.norm_genome')

} else if (norm.method == 'viral_read.count') {
    
  ## Viral read count method
  # sum up all the ORF counts

  normBase <- melt(gene.sample_count.sp, variable.name = 'sample', value.name = 'count')
  normBase <- normBase[,.(norm_base = sum(count)), by=sample]
  
  fig.dir <- paste0(fig.dir, '.norm_LoRTIA')
  
} else if (norm.method == 'total_read.count') {
  ## For total read count method

  # This is not the LoRTIA-processed read counts, so each read is counted twice because they are from dcDNA that contains both orientations. 
  # But since this is true for both the virus and the host, this will cause each ORF to be estimated down approx 2-fold but in an unbiased way.

  normBase <- fread('../rebasecall/read_count.fastq.txt')
  normBase[ , sample := gsub('.fastq', '', file)]
  normBase[ , norm_base := read_count]
  normBase <- normBase[,.(sample, norm_base)]
  
  fig.dir <- paste0(fig.dir, '.norm_total')
  
} else if (norm.method == 'host_read.count') {
  ## For total HOST read count method

  # This is not the LoRTIA-processed read counts, so each read is counted twice because they are from dcDNA that contains both orientations. 
  # But since this is true for both the virus and the host, this will cause each ORF to be estimated down approx 2-fold but in an unbiased way.
  # Currently this is the unmapped read count!

  normBase <- fread('../rebasecall/read_count.txt')
  #normBase[ , sample := gsub('.fastq', '', file)]
  #normBase <- normBase[,.(sample, unmapped)]
  normBase[ , norm_base := unmapped]
  normBase <- normBase[,.(sample, norm_base)]
  
  fig.dir <- paste0(fig.dir, '.norm_host')
  
} else if (norm.method == 'WO') {
  
  ## NO normalization of counts

  normBase <- fread('../rebasecall/read_count.txt')
  normBase[ , norm_base := 1]
  normBase <- normBase[,.(sample, norm_base)]
  
  fig.dir <- paste0(fig.dir, '.norm_Without')
  
}

try({ dir.create(fig.dir) })

##
normBase <- normBase[order(normBase$sample, colnames(orf.counts[-c(1:6)])), ]
stopifnot(all(normBase$sample == colnames(orf.counts[-c(1:6)]) ))

orf.perc <- merge(normBase, orf.sum, by='sample')
orf.perc[,read_count := read_count / norm_base]


#### TO-DO: 

## 1.) find which replicate is the farthest from the mean in each gene for each group -> done
## 2.) Carry out the moanin analyis using the current normalization 
## 3.) Check the overlap of Literature-based kinetic classes and de-novo clusters
```


```{r}
### Filtering of very-low coverage samples
samples_to_filt <- c('C6_2h_3', 'PC-12_8h_3')

orf.perc <- orf.perc[!orf.perc$sample %in% samples_to_filt, ]
metafilt <- metafilt[!metafilt$sample %in% samples_to_filt, ]

message(paste(samples_to_filt, ' was filtered out!', collapse = '\n') )

```


# 1. Counts based on TSS only (5` -ends of reads)

## Statistics of counts {.tabset}
```{r, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}

#### Keep those where the TSS was correct, regardless of TES
orf.sum.perc <- orf.perc
orf.sum.perc <- orf.sum.perc[!is.na(gene), ]

orf.sum.perc[,gene_cluster   := gene]

gene_kin.class <- na.omit(unique(orf.sum.perc[,.(gene, Kinetic_class)]))
gene_kin.class[,Kinetic_class := paste(Kinetic_class, collapse = '/'), by=.(gene)]
gene_kin.class <- unique(gene_kin.class[,.(gene, Kinetic_class)])

orf.sum.perc[,Kinetic_class := NULL]

#### Summarize the counts

## summarise counts on gene cluster
orf.sum.perc <- orf.sum.perc[,.(read_count=sum(read_count)), by=.(seqnames, strand, gene, norm_base, sample, group, hpi, Time, cell_line)]
orf.sum.perc <- merge(orf.sum.perc, gene_kin.class, by='gene', allow.cartesian=T, all=T)

### Statistics of gene cluster counts accross samples

## normalized counts
orf.perc.mean <- 
  orf.sum.perc %>% group_by(gene, Kinetic_class, hpi, Time, cell_line) %>%
  dplyr::summarise(mean=mean(read_count), sd=sd(read_count), varcoeff=sd(read_count) / mean(read_count), se=mean_se(read_count))
orf.perc.mean <- data.table(orf.perc.mean[,1:8], orf.perc.mean$se[])

orf.perc.mean[is.na(Kinetic_class), Kinetic_class := 'unknown']
orf.perc.mean[,Kinetic_class := factor(Kinetic_class, levels = c('IE', 'IE/E', 'E', 'IE/E/L', 'E/L', 'L', 'unknown'))]

## non-normalized counts
#orf.mean <- 
#  orf.sum %>% group_by(gene, hpi, Time, cell_line) %>%
#  dplyr::summarise(mean=mean(read_count), sd=sd(read_count), varcoeff=sd(read_count) / mean(read_count), se=mean_se(read_count))
#orf.mean <- data.table(orf.mean[,1:7], orf.mean$se[])



#### merge back summary with counts
orf.sum.perc <- merge(orf.sum.perc, orf.perc.mean, by=c('gene', 'Kinetic_class', 'hpi', 'Time', 'cell_line'))


```

### Difference from mean
```{r, fig.width = 9, fig.height = 6, eval=F}

#### Find which replicate is the farthest from the mean in each gene for each group

# find distance from mean
orf.sum.perc[,dist_from_mean := abs(mean - read_count)]

#
#orf.sum <- merge(orf.sum, orf.mean, by=c('gene', 'hpi', 'Time', 'cell_line'))
# find distance from mean
#orf.sum[,dist_from_mean := abs(mean - read_count)]


##
orf.sum.perc[ , farthest_from_mean := ifelse(dist_from_mean == max(dist_from_mean), T, F), by=.(group, gene)]
# NA, if all is true in each group beacuse then the distance is the same which means probably that the gene has zero counts in that group
orf.sum.perc[ , farthest_from_mean := if(sum(dist_from_mean) == 0) NA, by=.(group, gene)]
orf.sum.perc.tes[ , farthest_rep := ifelse(farthest_from_mean, gsub('.*_', '', sample), NA), by=.(sample, gene)]

plot.data  <- orf.sum.perc[!is.na(farthest_rep), .N, by=.(farthest_rep, hpi, Time, cell_line, group)][order(group)]

ggplot(
  plot.data,
  aes(x=hpi, y=N, fill=farthest_rep)) +
  geom_col(position = 'dodge2') +
  scale_fill_d3() +
  theme_bw() +
  facet_grid(rows=vars(cell_line))

```


### Varcoeff
```{r, fig.width = 20, fig.height = 8}

ggplot(
  orf.sum.perc[],
  aes(x=gene, y=varcoeff, fill=hpi)) +
  geom_col(position = 'dodge2') +
  scale_fill_d3() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) +
  facet_grid(rows=vars(cell_line))


```


```{r}
orf.perc.sum  <- orf.sum.perc
orf.perc.sum[,libtype := 'dcDNA']
orf.perc.sum[,sample_name := sample]
orf.perc.sum[,norm.method := norm.method]
```



## ORF kinetics, according to literature-based kinetic classes {.tabset}
Time-dependent normalized ORF counts (based on TSSs and TES clusters)

```{r}
### ORFs in samples calculated from TRS-B transcripts with and without leaders

max_ncol <- 10
row_multip <- 6
col_multip <- 4

breaks <- na.omit(unique(metafilt$Time)[order(unique(metafilt$Time))])
labels <- c('', '1', '', '2', '', '4', '6', '8', '12')


plotfun <- function(DT, 
                    geom1 = geom_pointrange(aes(ymin = mean-sd, ymax = mean+sd, colour = cell_line)),
                    geom2 = geom_smooth(aes(x=Time, y=mean, color=cell_line)),
                    ncol = 10, title = NULL, #"Normalized ORF counts", 
                    strip.backgr = pal_d3()(10)[2]) {
  
  if (nrow(DT) == 0) { 
    ggp <- NULL } else {   # try({ 
      
    # 
    gene_N <- length(unique(DT[,gene]))
    # class_freq <- DT[best_label == 5][,.N,by=class_col]
    
    #colid <- c('gene', class_col)
    #class_freq <- unique(DT[best_label == 5, ..colid])
    
    
    #  min(class_freq)
     # =='adaptive') {}
        
    ggp <- 
      ggplot(DT[
                                #  orf.perc.mean$hpi    != 'dRNA' 
                                #& orf.perc.mean$strand == '+'
                                , ], 
                     aes(Time, mean, color = 'cell_line')
                     #, add = 'loess'
                     ) + 
      geom1 +
      geom2 + 
      scale_color_manual(values = palette) +
      scale_y_continuous(labels = scales::percent_format()) +
      scale_x_continuous(breaks = breaks, labels=labels) +
      theme_ipsum() +
      theme(plot.margin = unit(c(1,1,1,1), 'mm')) +
      facet_wrap(~ Kinetic_class + gene, scales = 'free_y', ncol = ncol) +
      theme(strip.background = element_rect(fill = alpha(strip.backgr, 0.4)))
      # + ggtitle(title)
  
  if (gene_N < ncol) { 
    ggp <- cowplot::plot_grid(ggp, NULL, rel_widths = c(gene_N, ncol - gene_N ), nrow=1)
  }
  
  }   # })
  
  return(ggp)

}
```


```{r}
# Kinetic class frequency

kin_classes      <- unique(orf.perc.mean[,.(gene, Kinetic_class)])
all.class.freq   <- data.table(kin_classes)[,.N, by=.(Kinetic_class) ][order(Kinetic_class)]

my_pal <- pal_d3()(10)

mapfun <- function(i, DT, all.class.freq, palette = my_pal, ...) {
  kin_class <- all.class.freq[i, Kinetic_class]
  strip.background <- palette[i]
  plotfun(DT[Kinetic_class == kin_class],  strip.backgr = strip.background, ...)
}


ncol     <- ifelse(max(all.class.freq$N) < max_ncol, max(all.class.freq$N), max_ncol)
relh     <- ceiling(all.class.freq$N / ncol)  #c(6,1,1,1,2)
figw     <- ncol * col_multip
figh     <- nrow(all.class.freq) * row_multip

```


### Mean and SD with loess function
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- purrr::map(seq_along(relh), mapfun, orf.perc.mean, all.class.freq, ncol=ncol)

ggp <- cowplot::plot_grid(plotlist = ggp, ncol = 1, rel_heights = relh, align = 'v', axis=c('tlrb'))

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method,  '_Literature.Clusters', '_mean.sd.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp
```

### Mean and SD with linear connection
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- purrr::map(seq_along(relh), mapfun, orf.perc.mean, all.class.freq, ncol=ncol, 
                  geom1 = geom_pointrange(aes(ymin = mean-sd, ymax = mean+sd, colour = cell_line)),
                  geom2 = geom_line(aes(x=Time, y=mean, color=cell_line)))

ggp <- cowplot::plot_grid(plotlist = ggp, ncol = 1, rel_heights = relh)


ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Literature.Clusters', '_mean.sd.linear.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp
```

### Each replicate with loess
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- purrr::map(seq_along(relh), mapfun, orf.sum.perc, all.class.freq, ncol=ncol, 
                  geom1 = geom_point( aes(x=Time, y=read_count, color=cell_line)),
                  geom2 = geom_smooth(aes(x=Time, y=read_count, color=cell_line)) )

ggp <- cowplot::plot_grid(plotlist = ggp, ncol = 1, rel_heights = relh)


ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Literature.Clusters', '_replicates.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp
```




## Combine Kinetic classes

sum up all the (normalized) gene counts for each kinetic class in each sample (replicate)
then take the mean and SD of the sums for each hpi and cell line

```{r}
### Summarise based on literature kinetic class for each sample
kin.class.sum  <- orf.sum.perc[,.(sum_read_count=sum(read_count)),  
                               by=.(seqnames, Kinetic_class, hpi, Time, cell_line, group, sample, norm_base)]

### take the mean for each hpi and cell line (without sample)
kin.class.sum  <- kin.class.sum[, .(mean_read_count=mean(sum_read_count), sd_read_count=sd(sum_read_count)),
                                by=.(seqnames, Kinetic_class, hpi, Time, cell_line, group, norm_base)]
                                

```

```{r}
figw <- 16
figh <- 8
ncol <- 10
```


### Literature clusters
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- 
      ggplot(kin.class.sum,
             aes(Time, mean_read_count, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_point() +
      geom_smooth() + 
      scale_color_manual(values = palette[]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      scale_y_continuous(labels = scales::percent_format()) +
      scale_x_continuous(breaks = breaks, labels=labels) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      # facet_nested(rows=vars(norm.method), cols=vars(labels), scales = "free_y") ## cols=vars(norm.method)
      facet_wrap(~ Kinetic_class, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.Literature.Clusters.', '_points.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)

```



## Literature Clusters and Selected Genes {.tabset}
```{r}
### Summarise based on literature kinetic class
select.genes.class.sum <- 
  orf.perc.sum[!gene %in% c('US1', 'EP0', 'IE180'),
               .(sum_read_count=sum(read_count)), 
               by=.(Kinetic_class, hpi, Time, cell_line, seqnames, norm_base, sample, group)]

select.genes.class.sum <- rbind(
  select.genes.class.sum,
  orf.perc.sum[gene %in% c('US1', 'EP0', 'IE180'),
               .(sum_read_count=sum(read_count)), 
               by=.(gene, hpi, Time, cell_line, seqnames, norm_base, sample, group)],
  fill = T
)

select.genes.class.sum[gene %in% c('US1', 'EP0', 'IE180'), Kinetic_class := gene]


### take the mean for each hpi and cell line (without sample)
select.genes.class.cell.lines  <- select.genes.class.sum[, .(mean_cell_line=mean(sum_read_count), sd_cell_line=sd(sum_read_count)),
                                by=.(seqnames, Kinetic_class, hpi, Time, cell_line, group)]



#select.genes.class.sum[,mean_cell_line := mean(mean_read_count), by=.(seqnames, Kinetic_class, hpi, Time, cell_line, group)]

#select.genes.class.sum[,sd_cell_line   := sd(mean_read_count), by=.(seqnames, Kinetic_class, hpi, Time, cell_line, group)]


#select.genes.class.cell.lines <- unique(select.genes.class.sum[, .(mean_cell_line, sd_cell_line, seqnames, Kinetic_class, hpi, Time, cell_line, group)] )

```


```{r}
figw <- 16
figh <- 10
ncol <- 10
```


### Reps with loess and confidence
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ncol=3

ggp <- 
      ggplot(select.genes.class.sum,
             aes(Time, sum_read_count, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_point() +
      geom_smooth() + 
      scale_color_manual(values = palette[]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      scale_y_continuous(labels = scales::percent_format()) +
      scale_x_continuous(breaks = breaks, labels=labels) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      #facet_nested(cols=vars(Kinetic_class), scales = "free_y") ## cols=vars(norm.method), rows=vars(norm.method), 
      facet_wrap(~ Kinetic_class, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.Literature.Clusters.SELECT', '_points.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)
```

### Means with SD and linear connection
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- 
      ggplot(select.genes.class.cell.lines,
             aes(Time, mean_cell_line, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_pointrange(aes(Time, ymin=mean_cell_line - sd_cell_line, ymax=mean_cell_line + sd_cell_line, color = cell_line)) +    
      geom_point() +
      geom_line(aes(group = cell_line)) + 
      scale_color_manual(values = palette[]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      scale_y_continuous(labels = scales::percent_format()) +
      scale_x_continuous(breaks = breaks, labels=labels) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      #facet_nested(cols=vars(Kinetic_class), scales = "free_y") ## cols=vars(norm.method), rows=vars(norm.method),
      facet_wrap(~ Kinetic_class, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.Literature.Clusters.SELECT', '_mean.SD.linear.jpg'), ggp, width = figw, height = figh, limitsize = F)
```



## Gene-pair ratios {.tabset}

These pairs were analyzed: 
ie180/us1, ie180/ep0, us1/ep0, noir1/noir2, noir1/ie180, noir2/ie180

```{r, fig.show='hold',  fig.width = 10, fig.height = 7, eval=T}
ab.US1    <- setnames(orf.perc.sum[gene == 'US1',    .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'US1')

ab.IE180  <- setnames(orf.perc.sum[gene == 'IE180',  .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'IE180')

ab.EP0    <- setnames(orf.perc.sum[gene == 'EP0',    .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'EP0')

ab.noir   <- setnames(orf.perc.sum[gene == 'NOIR',   .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'NOIR')

ab.noir2  <- setnames(orf.perc.sum[gene == 'NOIR-2', .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'NOIR-2')


ab.merged <- merge(ab.US1,    ab.IE180,  by=c('sample', 'hpi', 'Time', 'cell_line', 'group', 'norm.method'))
ab.merged <- merge(ab.merged, ab.EP0,    by=c('sample', 'hpi', 'Time', 'cell_line', 'group', 'norm.method'))
ab.merged <- merge(ab.merged, ab.noir,   by=c('sample', 'hpi', 'Time', 'cell_line', 'group', 'norm.method'))
#ab.merged <- merge(ab.merged, ab.noir2,  by=c('sample', 'hpi', 'Time', 'cell_line', 'norm.method'))

ab.merged[, IE180_US1   :=  IE180   /  US1]
ab.merged[, IE180_EP0   :=  IE180   /  EP0]
ab.merged[, US1_EP0     :=  US1     /  EP0]
#ab.merged[, NOIR1_NOIR2 :=  NOIR    / `NOIR-2`]
ab.merged[, NOIR_IE180  :=  NOIR    /  IE180]
#ab.merged[, NOIR2_IE180 := `NOIR-2` /  IE180]

#ab.merged[, US1_IE180   :=  US1     / IE180     ]
#ab.merged[, EP0_IE180   :=  EP0     / IE180     ]
#ab.merged[, EP0_US1     :=  EP0     / US1       ]
#ab.merged[, NOIR2_NOIR1 := `NOIR-2` / NOIR      ]
ab.merged[, IE180_NOIR  :=  IE180   / NOIR      ]
#ab.merged[, IE180_NOIR2 :=  IE180   / `NOIR-2`  ]

gpairs <- c(
'IE180_US1' , 
'IE180_EP0'  , 
'US1_EP0'     ,
'NOIR1_NOIR2' ,
'NOIR_IE180'  ,
'NOIR2_IE180' ,

'US1_IE180'   ,
'EP0_IE180'   ,
'EP0_US1'     ,
'NOIR2_NOIR1' ,
'IE180_NOIR'  ,
'IE180_NOIR2' 
)
gpairs <- gsub('_', '/', gpairs)

ab.merged.gt <- ab.merged#[,-c(7:11)]

ab.merged.gt <- melt(ab.merged.gt, variable.name = 'gene_pair', value.name = 'ratio', id.vars = c(1:6))

ab.merged.gt$gene_pair <- gsub('_', '/', ab.merged.gt$gene_pair)

ab.merged.gt <- ab.merged.gt[gene_pair %in% gpairs, ]

ab.merged.gt[,gene_pair := factor(gene_pair, levels = gpairs)]

ab.merged.gt[,mean := mean(ratio),   by=.(hpi, Time, cell_line, group, norm.method, gene_pair)]
ab.merged.gt[,sd   := sd  (ratio),   by=.(hpi, Time, cell_line, group, norm.method, gene_pair)]
ab.merged.gt[,sum  := sum (ratio),   by=.(hpi, Time, cell_line, group, norm.method, gene_pair)]



plot_data <- ab.merged.gt[norm.method == 'viral_read.count'
                 # norm.method == 'host_read.count' 
                 , ]
```


```{r}
figw <- 16
figh <- 10
ncol <- 10
```


### Means with SD and linear connection
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}
ggmean <- ggplot(plot_data) + 
  #geom_point(aes(Time, ratio, color=cell_line)) +
  geom_line(aes(Time, mean, color=cell_line, group=cell_line)) +
  geom_pointrange(aes(Time, mean, color=cell_line, group=cell_line, ymin=mean-sd, ymax=mean+sd)) +
  scale_color_manual(values = palette[]) +
  scale_y_continuous(name='ratio') +
  scale_x_continuous(breaks = breaks, labels=labels) +
  theme_ipsum() +
  #facet_nested(rows = vars(norm.method), cols=vars(Time))
  facet_wrap(~gene_pair, scales = 'free', ncol=3)


ggmean
ggsave(paste0(fig.dir, '/gene_ratios_mean.jpg'), width = 10, height = 7)
```



### Reps with loess and confidence
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggrep <- ggplot(plot_data[# norm.method == 'viral_read.count'
                 # norm.method == 'host_read.count' 
                 , ]) + 
  geom_point(aes(Time, ratio, color=cell_line)) +
  geom_smooth(aes(Time, ratio, color=cell_line, group=cell_line)) +
  #geom_pointrange(aes(Time, mean, color=cell_line, group=cell_line, ymin=mean-sd, ymax=mean+sd)) +
  scale_color_manual(values = palette[]) +
  scale_y_continuous(name='ratio') +
  scale_x_continuous(breaks = breaks, labels=labels) +
  theme_ipsum() +
  #facet_nested(rows = vars(norm.method), cols=vars(Time))
  facet_wrap(~gene_pair, scales = 'free', ncol=3)


ggrep
ggsave(paste0(fig.dir, '/gene_ratios_rep.jpg'), width = 10, height = 7)
```


### Sums with linear connection
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}
ggsum <- ggplot(plot_data) + 
  geom_point(aes(Time, sum, color=cell_line)) +
  geom_line(aes(Time, sum, color=cell_line, group=cell_line)) +
  #geom_pointrange(aes(Time, mean, color=cell_line, group=cell_line, ymin=mean-sd, ymax=mean+sd)) +
  scale_color_manual(values = palette[]) +
  scale_y_continuous(name='ratio') +
  scale_x_continuous(breaks = breaks, labels=labels) +
  theme_ipsum() +
  #facet_nested(rows = vars(norm.method), cols=vars(Time))
  facet_wrap(~gene_pair, scales = 'free', ncol=3)


ggsum
ggsave(paste0(fig.dir, '/gene_ratios_sum.jpg'), width = 10, height = 7)

```


```{r}
## Export data
if (writetables) {
  fwrite(orf.sum.perc,       paste0(fig.dir, '/', norm.method, '_', EndType, '_counts.norm.tsv'), sep = '\t')
  fwrite(orf.perc.mean,      paste0(fig.dir, '/', norm.method, '_', EndType, '_abund.mean.tsv'), sep = '\t')
  #fwrite(DE_results,         paste0(fig.dir, '/', norm.method, '_', EndType, '_DE_results.tsv'), row.names=T, sep = '\t')
  #fwrite(DE_cluster_results, paste0(fig.dir, '/', norm.method, '_', EndType, '_moanin_DE_results.tsv'), sep = '\t')

}
```


```{r bib, include=FALSE}
# KEEP THIS AT THE END OF THE DOCUMENT TO GENERATE A LOCAL bib FILE FOR PKGS USED
knitr::write_bib(sub("^package:", "", grep("package", search(), value=TRUE)), file='skeleton.bib')
```