---
title: "PRV - 4 cell lines: TSS dynamics"
author: Balázs Kakuk
output: 
  html_document:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: false
---

```{css, echo=FALSE}
    body .main-container {
      max-width: 90% !important;
      width: 90% !important;
    }
    body {
      max-width: 90% !important;
      margin-left: auto;
      margin-right: auto;
    }
```


```{r include=FALSE}
## hrbrthemes::ipsum:

knitr::opts_chunk$set(fig.retina=2, echo = FALSE, fig.align = 'center', message = T, warning = F)
```

```{r ipsum_setup, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE, include=FALSE}
library(prettydoc)
library(hrbrthemes, quietly = T)
#library(GenomicFeatures, quietly = T)
library(DESeq2, quietly = T)
library(ggpubr, quietly = T)
#library(misc, quietly = T)
library(dplyr, quietly = T)
#library(tidyverse, quietly = T)
library(ggsci, quietly = T)
#library(Gviz, quietly = T)
library(dplyr, quietly = T)
library(tidyr)
library(scales)
#library(plyr)
#library(moanin, quietly = T)
library(grid, quietly = T)
library(gridExtra, quietly = T) 
library(ggpubr, quietly = T)
library(data.table)
library(ggh4x)
library(knitr)
library(formattable)



### Project name and output directory
outdir  <- 'PRV-MDBIO-4cell'; try({ dir.create(outdir) })

#### Load configuration file
config <- readRDS(paste0(outdir, '_config.rds'))

####
source('_WF.part0.R')

#### Settings ####

## Main
#save.data <- 'PRV.rebase.RData'
EndType <- 'TSS'

#project_config  <- fread('project_config.txt')
#outdir  <- project_config$outdir
res.dir <- outdir; try({ dir.create(res.dir) })
fig.dir <- paste0(res.dir, '/', EndType, '_abund'); # try({ dir.create(fig.dir) })## save plots to this directory

## Miscallenaous
## Colors
colorvec <- na.omit(c(
  colorvec[c(1:5,15,13,10,24,7,26,32,33)],
  colorvec[-c(1:5,15,13,10,24,7,26,32,33)]))

colorvec <- colorvec[-c(9,13)]


colorvec <- colorvec[-c(10,13,17,18)]

colorvec <- colorvec[c(1:2,14,3:length(colorvec))]

palette <- colorvec  #pal_npg()(10)

##
writetables <- T
load.prev   <- F



## Filter metadata for those samples that had at least 1 viral read
readcounts <- fread(paste0(outdir, '/readcounts.tsv'), header = T)
metafilt   <- metafilt[metafilt$sample %in% readcounts$sample, ]


#### Normalization method
norm.method <- 'viral_read.count'
# Normalization before or after filtering for counts that didn't overlap with the other end?
norm_before_filtering <- F

```


# Project info
Samples: dcDNA Seq  
pychopper: no  
mapping version: v6  
LoRTIA: yes (stranded only output)  

Filtering samples:  
- C6_2h_3  
- PC-12_8h_3  
- C6_0.5h_1  
- C6_0.5h_2  
- C6_0.5h_3  
- PC-12_0.5h_3  
- PK-15_2h3

Normalization: Viral read count

## Only those reads were counted, where LoRTIA found *5-prime* adapter sequences !


```{r, eval=T}
#### Import TSS counts

gene.sample_count.sp <- fread(paste0(res.dir, '/gene.sample.count.sp.tsv'), na.strings = '')


#### !!! Consider adapters!
       if (EndType == 'TES')  { adapters <- 'prime3' 
} else if (EndType == 'TSS')    { adapters <- 'prime5' 
} else if (EndType == 'TSS.TES')  { adapters <- 'both' } ## OR: 'either'}

source('Gene.counts.R')

#gene.sample_count    <- TR.ref.sum[,.(count=sum(read_count)), by=.(seqnames, strand, gene_cluster, gene, Kinetic_class, sample)]
#gene.sample_count.sp <- dcast(gene.sample_count, seqnames + strand + gene_cluster + gene + Kinetic_class ~ sample)
####

orf.sum <- gene.sample_count.sp

## Filter out multi-copied genes
orf.sum <- orf.sum[!gene %in% c('IE180_2', 'US1_2', 'NOIR2', 'NOIR2-2'),]
orf.sum <- orf.sum[!gene_cluster %in% c('IE180_2', 'US1_2', 'NOIR2', 'NOIR2-2'),]

## Format columns
#orf.sum <- orf.sum[,!c("seqnames", "gene_cluster"), with = FALSE]
#setnames(orf.sum, old=c('Kinetic_class'), new=c('ID'))
orf.sum[,ID := gene]

# Reordering columns to make new_column the first column
setcolorder(orf.sum, c("ID", setdiff(names(orf.sum), "ID")))

## Factorize
orf.sum$gene          <- factor(orf.sum$gene,  levels=ORFs)
orf.sum$ID            <- factor(orf.sum$ID,    levels=ORFs)
orf.sum$Kinetic_class <- factor(orf.sum$Kinetic_class, levels=c('IE', 'E', 'L'))

## Order
orf.sum        <- orf.sum[order(Kinetic_class, gene),]

##
orf.counts     <- as.data.frame(orf.sum)

orf.sum        <- merge(gather(orf.sum, sample, read_count, -c(1:6)), metafilt[,metacols], by='sample')
setDT(orf.sum)
```


```{r, eval=T}
#### Normalization Bases

 if (norm.method == 'average_genome') {
  
  ## Average genome norm method
  norm_cov_summary <- fread('norm.cov.summary.tsv')
  normBase <- norm_cov_summary[,.(sample, average_coverage)]
  colnames(normBase)[2] <- 'norm_base'
  
  fig.dir <- paste0(fig.dir, '.norm_genome')

} else if (norm.method == 'viral_read.count') {
    
  ## Viral read count method
  # sum up all the ORF counts

  #normBase <- melt(gene.sample_count.sp, variable.name = 'sample', value.name = 'count')
  #normBase <- normBase[,.(norm_base = sum(count)), by=sample]
  
  # After removing counts for the multi-copy genes
  normBase <- orf.sum[,.(norm_base = sum(read_count)), by=sample]
  
  fig.dir <- paste0(fig.dir, '.norm_LoRTIA')
  
} else if (norm.method == 'total_read.count') {
  ## For total read count method

  # This is not the LoRTIA-processed read counts, so each read is counted twice because they are from dcDNA that contains both orientations. 
  # But since this is true for both the virus and the host, this will cause each ORF to be estimated down approx 2-fold but in an unbiased way.

  normBase <- fread('../rebasecall/read_count.fastq.txt')
  normBase[ , sample := gsub('.fastq', '', file)]
  normBase[ , norm_base := read_count]
  normBase <- normBase[,.(sample, norm_base)]
  
  fig.dir <- paste0(fig.dir, '.norm_total')
  
} else if (norm.method == 'host_read.count') {
  ## For total HOST read count method

  # This is not the LoRTIA-processed read counts, so each read is counted twice because they are from dcDNA that contains both orientations. 
  # But since this is true for both the virus and the host, this will cause each ORF to be estimated down approx 2-fold but in an unbiased way.
  # Currently this is the unmapped read count!

  normBase <- fread('../rebasecall/read_count.txt')
  #normBase[ , sample := gsub('.fastq', '', file)]
  #normBase <- normBase[,.(sample, unmapped)]
  normBase[ , norm_base := unmapped]
  normBase <- normBase[,.(sample, norm_base)]
  
  fig.dir <- paste0(fig.dir, '.norm_host')
  
} else if (norm.method == 'WO') {
  
  ## NO normalization of counts

  normBase <- fread('../rebasecall/read_count.txt')
  normBase[ , norm_base := 1]
  normBase <- normBase[,.(sample, norm_base)]
  
  fig.dir <- paste0(fig.dir, '.norm_Without')
  
}

try({ dir.create(fig.dir) })

##
#normBase <- normBase[order(normBase$sample, colnames(orf.counts[-c(1:6)])), ]
#stopifnot(all(normBase$sample == colnames(orf.counts[-c(1:6)]) ))

orf.perc <- merge(normBase, orf.sum, by='sample')
#orf.perc[, read_count := read_count / norm_base]
```


```{r, eval=T}

orf.sum.perc <- orf.perc


#### Normalization: Before Filtering
if (norm_before_filtering & norm.method == 'viral_read.count') { 
  orf.sum.perc[, read_count := read_count / norm_base]
}


###### Filter for counts that have either or both ends overlap with TSS or TES
if (EndType == 'TSS') {

  #### Keep those where the TSS was correct, regardless of TES
  orf.sum.perc <- orf.sum.perc[!is.na(gene), ]
  
  orf.sum.perc[,gene_cluster   := gene]
  
} else if (EndType == 'TES') {
  
  #### Keep those where the TES was correct, regardless of TSS
  orf.sum.perc <- orf.sum.perc[!is.na(gene_cluster), ]
  
  orf.sum.perc[,gene   := gene_cluster]
  
} else if (EndType == 'TSS.TES') {
  
  #### Keep those where both the TSS and TES were correct
  orf.sum.perc <- orf.sum.perc[!is.na(gene_cluster) & !is.na(gene), ]
  
  orf.sum.perc[,gene_cluster   := gene]
  
}


#### Normalization: After Filtering
if (!norm_before_filtering  & norm.method == 'viral_read.count') {
  orf.sum.perc[, norm_base  := sum(read_count), by = sample]
  orf.sum.perc[, read_count := read_count / norm_base]
}

## check normalizations
#orf.sum.perc[,.(sum = sum(read_count)), by=sample]

```


```{r}

## Load previous results ?

if (load.prev) {
  
  fig.dir <- paste0(fig.dir, '.norm_LoRTIA')

  orf.sum.perc  <- fread( paste0(fig.dir, '/', norm.method, '_', EndType, '_abund.tsv') )
  
  orf.sum.perc[, mean     := NULL]
  orf.sum.perc[, varcoeff := NULL]
  orf.sum.perc[, sd       := NULL]
  orf.sum.perc[, cluster  := NULL]
  #orf.perc.mean <- fread( paste0(fig.dir, '/', norm.method, '_', EndType, '_abund.mean.tsv')  )
  #setnames(orf.sum.perc, c('mean.y', 'sd.y', 'varcoeff.y'), c('mean', 'sd', 'varcoeff'))
  
  metafilt   <- metafilt[metafilt$sample %in% orf.sum.perc$sample, ]
  
}
```


```{r, eval=T, message=T}

#min_read_count <- 10


### Filtering of very-low coverage samples
samples_to_filt <- c('C6_2h_3', 'PC-12_8h_3', 'C6_0.5h_1', 'C6_0.5h_2', 'PC-12_0.5h_3', #'PK-15_2h_3',
                     grep('dRNA', metafilt$sample, value = T))

orf.sum.perc  <- orf.sum.perc[!orf.sum.perc$sample %in% samples_to_filt, ]

metafilt <- metafilt[!metafilt$sample %in% samples_to_filt, ]

message(paste(samples_to_filt, ' was filtered out!', collapse = '\n') )

```


# 1. Counts based on TSS only (5` -ends of reads)

## Statistics of counts {.tabset}
```{r, eval = T}


#### Format Kinetic Class
gene_kin.class <- na.omit(unique(orf.sum.perc[,.(gene, Kinetic_class)]))
gene_kin.class[,Kinetic_class := paste(Kinetic_class, collapse = '/'), by=.(gene)]
gene_kin.class <- unique(gene_kin.class[,.(gene, Kinetic_class)])

orf.sum.perc[,Kinetic_class := NULL]
orf.sum.perc <- merge(orf.sum.perc, gene_kin.class, by='gene', all.x=T)

orf.sum.perc[is.na(Kinetic_class), Kinetic_class := 'unknown']
orf.sum.perc[,Kinetic_class := factor(Kinetic_class, levels = c('IE', 'IE/E', 'E', 'IE/E/L', 'E/L', 'L', 'unknown'))]



#### Summarize the counts on gene
orf.sum.perc <- orf.sum.perc[,.(read_count=sum(read_count)), by=.(seqnames, strand, gene, norm_base, sample, group, hpi, Time, cell_line, Kinetic_class)]


#### Statistics of gene cluster counts across replicates
orf.perc.mean <- orf.sum.perc[, .(mean=mean(read_count), sd=sd(read_count), varcoeff=sd(read_count) / mean(read_count)), 
                              by=.(gene, Kinetic_class, hpi, Time, cell_line, group)]

#### Merge back summary with counts
orf.sum.perc <- merge(orf.sum.perc, orf.perc.mean, 
                      by=c('gene', 'Kinetic_class', 'hpi', 'Time', 'cell_line', 'group'))


```



```{r}


orf.sum.sp     <- dcast.data.table(orf.sum.perc, gene~sample, value.var = 'read_count')

orf.sum.sp.mat <- as.matrix(data.frame(orf.sum.sp[,-1], row.names = orf.sum.sp$gene))
colnames(orf.sum.sp.mat) <- colnames(orf.sum.sp)[-1]

colData <- data.frame(metafilt, row.names = metafilt$sample)


orf.sum.sp.mat <- orf.sum.sp.mat[, rownames(colData)]

colnames(orf.sum.sp.mat)[which(!colnames(orf.sum.sp.mat) %in% rownames(colData))]
  

if (norm.method == 'VST') {  
  deseq <- DESeqDataSetFromMatrix(orf.sum.sp.mat, colData = colData, ~ hpi)
  deseq <- estimateSizeFactors(deseq, type =  "poscounts")
  deseq <- DESeq(deseq)
  results(deseq)
  
  vsd      <- varianceStabilizingTransformation(deseq)
  vsd_mat  <- assay(vsd)[,metafilt$sample]
  vsd_data <- data.table(gene = rownames(vsd_mat), vsd_mat)

  vsd_melt <- melt.data.table(vsd_data, 
      id.vars = 'gene',
      value.name = 'norm_read_count', 
      variable.name = 'sample')
  #vsd_melt <- merge(vsd_melt, metafilt, by='sample')

  orf.sum.percx <- merge(orf.sum.perc, 
                        vsd_melt,
                        by=c('gene', 'sample'))
  if(nrow(orf.sum.percx) == nrow(orf.sum.perc)) {
    orf.sum.perc <- orf.sum.percx
    rm(orf.sum.percx)
  } else { stop() }
  
  orf.sum.sp.mat <- vsd_mat
  
}
```


```{r, eval=T}


### Statistics of gene cluster counts accross samples
orf.perc.mean <- 
  orf.sum.perc[,.(mean=mean(read_count), sd=sd(read_count), varcoeff=sd(read_count) / mean(read_count)), 
               by=.(gene, Kinetic_class, hpi, Time, cell_line, group)]


#### merge back summary with counts
orf.sum.perc <- merge(orf.sum.perc, orf.perc.mean, by=c('gene', 'Kinetic_class', 'hpi', 'Time', 'cell_line', 'group'))


### spreaded means
orf.perc.mean.sp  <- dcast.data.table(orf.perc.mean, gene~group, value.var = 'mean')

orf.perc.mean.mat <- as.matrix(data.frame(orf.perc.mean.sp[,-1], row.names = orf.perc.mean.sp$gene))


#orf.sum.perc[,.(sum = sum(read_count)), by=sample]
```


### Difference from mean
```{r, fig.width = 9, fig.height = 6, eval=F}

#### Find which replicate is the farthest from the mean in each gene for each group

# find distance from mean
orf.sum.perc[,dist_from_mean := abs(mean - read_count)]

#
#orf.sum <- merge(orf.sum, orf.mean, by=c('gene', 'hpi', 'Time', 'cell_line'))
# find distance from mean
#orf.sum[,dist_from_mean := abs(mean - read_count)]


##
orf.sum.perc[ , farthest_from_mean := ifelse(dist_from_mean == max(dist_from_mean), T, F), by=.(group, gene)]
# NA, if all is true in each group beacuse then the distance is the same which means probably that the gene has zero counts in that group
orf.sum.perc[ , farthest_from_mean := if(sum(dist_from_mean) == 0) NA, by=.(group, gene)]
orf.sum.perc.tes[ , farthest_rep := ifelse(farthest_from_mean, gsub('.*_', '', sample), NA), by=.(sample, gene)]

plot.data  <- orf.sum.perc[!is.na(farthest_rep), .N, by=.(farthest_rep, hpi, Time, cell_line, group)][order(group)]

ggplot(
  plot.data,
  aes(x=hpi, y=N, fill=farthest_rep)) +
  geom_col(position = 'dodge2') +
  scale_fill_d3() +
  theme_bw() +
  facet_grid(rows=vars(cell_line))

```


### Varcoeff
```{r, fig.width = 20, fig.height = 8, eval=F}

ggplot(
  orf.sum.perc[],
  aes(x=gene, y=varcoeff, fill=hpi)) +
  geom_col(position = 'dodge2') +
  scale_fill_d3() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) +
  facet_grid(rows=vars(cell_line))


```


```{r}
orf.perc.sum  <- orf.sum.perc
orf.perc.sum[,libtype := 'dcDNA']
orf.perc.sum[,sample_name := sample]
orf.perc.sum[,norm.method := norm.method]
```


## Clustering of PK-15 Samples

```{r}
method.hclust = "complete"
method.dist   = "uncentered"
```

```{r, fig.width=16, fig.height=10}
countData <- orf.sum.sp.mat[, metafilt$sample[metafilt$cell_line == 'PK-15'] ]


keep <- apply(countData, 1, function(x) var(x) > 0)

message('Filtering genes with very low counts:\n', paste(names(keep[keep == F]), collapse = '\n'))

countData <- countData[keep, ]

## Transform?
#countData <- log2(countData + 1e-5)

```

```{r}

source('pvclust.samples.R')

```



## Clustering of genes into de-novo clusters using hierarchial clustering based on PK-15

### Cluster memberships

```{r}

## use pvclust for statistical support
source('pvclust.R')

## use simple h-clustering
#source('hcluster.R')

```

### Optimal number of clusters

```{r, fig.width=16, fig.height=10}

source('hclust.wss.R')

```


### Finding the optimal number of clusters to cut the tree at and calculate {.tabset}
```{r}
sil_width_thresh   <- 0
scaled_dist_thresh <- 20
```


#### Clustering statistics: Cluster number: 5
```{r, fig.height=12, fig.width=12}

cluster_num <- 5

source('pvclust.cut.R')

source('cluster.eval.R')

ggclust_stats
```

```{r, fig.height=12, fig.width=12, echo=FALSE}

unclustered <- unique(sil_df[sil_width < sil_width_thresh | scaled_dist >= scaled_dist_thresh, gene])
message('Unclustered Genes, based on a Silhouette threshold of ', sil_width_thresh, ' and a Scaled Distance of ',scaled_dist_thresh, ': \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)

unclustered <- unique(merged_stats[au < 0.95, gene])
message('Unclustered Genes, based on AU threshold of 0.95: \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)
```

#### Clustering statistics: Cluster number: 6

```{r, fig.height=12, fig.width=12}

cluster_num <- 6

source('pvclust.cut.R')

source('cluster.eval.R')

ggclust_stats
```

```{r, fig.height=12, fig.width=12, echo=FALSE}

unclustered <- unique(sil_df[sil_width < sil_width_thresh | scaled_dist >= scaled_dist_thresh, gene])
message('Unclustered Genes, based on a Silhouette threshold of ', sil_width_thresh, ' and a Scaled Distance of ',scaled_dist_thresh, ': \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)

unclustered <- unique(merged_stats[au < 0.95, gene])
message('Unclustered Genes, based on AU threshold of 0.95: \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)
```


#### Clustering statistics: Cluster number: 7
```{r, fig.height=12, fig.width=12}

cluster_num <- 7

source('pvclust.cut.R')

source('cluster.eval.R')

ggclust_stats
```

```{r, fig.height=12, fig.width=12, echo=FALSE}

unclustered <- unique(sil_df[sil_width < sil_width_thresh | scaled_dist >= scaled_dist_thresh, gene])
message('Unclustered Genes, based on a Silhouette threshold of ', sil_width_thresh, ' and a Scaled Distance of ',scaled_dist_thresh, ': \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)

unclustered <- unique(merged_stats[au < 0.95, gene])
message('Unclustered Genes, based on AU threshold of 0.95: \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)
```

#### Clustering statistics: Cluster number: 8
```{r, fig.height=12, fig.width=12}

cluster_num <- 8

source('pvclust.cut.R')

source('cluster.eval.R')

ggclust_stats
```

```{r, fig.height=12, fig.width=12, echo=FALSE}

unclustered <- unique(sil_df[sil_width < sil_width_thresh | scaled_dist >= scaled_dist_thresh, gene])
message('Unclustered Genes, based on a Silhouette threshold of ', sil_width_thresh, ' and a Scaled Distance of ',scaled_dist_thresh, ': \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)

unclustered <- unique(merged_stats[au < 0.95, gene])
message('Unclustered Genes, based on AU threshold of 0.95: \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)
```


#### Clustering statistics: Cluster number: 10
```{r, fig.height=12, fig.width=12}

cluster_num <- 10

source('pvclust.cut.R')

source('cluster.eval.R')

ggclust_stats
```

```{r, fig.height=12, fig.width=12, echo=FALSE}

unclustered <- unique(sil_df[sil_width < sil_width_thresh | scaled_dist >= scaled_dist_thresh, gene])
message('Unclustered Genes, based on a Silhouette threshold of ', sil_width_thresh, ' and a Scaled Distance of ',scaled_dist_thresh, ': \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)

unclustered <- unique(merged_stats[au < 0.95, gene])
message('Unclustered Genes, based on AU threshold of 0.95: \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)
```


#### Clustering statistics: Cluster number: 12
```{r, fig.height=12, fig.width=12}
cluster_num <- 12

source('pvclust.cut.R')

source('cluster.eval.R')

ggclust_stats
```

```{r, fig.height=12, fig.width=12, echo=FALSE}

unclustered <- unique(sil_df[sil_width < sil_width_thresh | scaled_dist >= scaled_dist_thresh, gene])
message('Unclustered Genes, based on a Silhouette threshold of ', sil_width_thresh, ' and a Scaled Distance of ',scaled_dist_thresh, ': \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)

unclustered <- unique(merged_stats[au < 0.95, gene])
message('Unclustered Genes, based on AU threshold of 0.95: \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)
```



### Selected number of clusters: 12
```{r, fig.height=12, fig.width=12}

cluster_num <- 12

source('pvclust.cut.R')

source('cluster.eval.R')

ggclust_stats
```

```{r, fig.height=12, fig.width=12, echo=FALSE}

unclustered <- unique(sil_df[sil_width < sil_width_thresh | scaled_dist >= scaled_dist_thresh, gene])
message('Unclustered Genes, based on a Silhouette threshold of ', sil_width_thresh, ' and a Scaled Distance of ',scaled_dist_thresh, ': \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)

unclustered <- unique(merged_stats[au < 0.95, gene])
message('Unclustered Genes, based on AU threshold of 0.95: \n',
        paste(unclustered, collapse = '\n') )
#length(unclustered)
```



### Remove genes from clusters ?
```{r, message = T, eval=T}
## Remove based on cluster significance
cluster_dt[, cluster := paste0('cluster_', cluster)]

unclustered <- unique(merged_stats[au < 0.95, gene])
unclustered
length(unclustered)


rm.unclustered <- F

if (rm.unclustered) {
  
  message('The following genes were removed from their corresponding clusters 
           as their silhouette widths were lower than 0.1 and their distance from the group centroid was higher than 0.5: \n', 
          paste(unclustered, collapse = '\n'))
  cluster_dt[gene %in% unclustered, cluster := 'unclustered'] 
  
} else {
  
  message('No genes were removed from their corresponding clusters.')
  
}
#
merged_stats[,.N,cluster][order(N, decreasing = T)]

```

```{r, message = T, eval=F}
### Remove based on silhouettes --> Select one option!
cluster_dt[, cluster := paste0('cluster_', cluster)]

unclustered <- unique(sil_df[sil_width < sil_width_thresh | scaled_dist >= scaled_dist_thresh, gene])
unclustered
length(unclustered)


rm.unclustered <- T

if (rm.unclustered) {
  
  message('The following genes were removed from their corresponding clusters 
           as their silhouette widths were lower than ', sil_width_thresh, ' and their scaled distance from the group centroid was higher than ', scaled_dist_thresh, ': \n', 
          paste(unclustered, collapse = '\n'))
  cluster_dt[gene %in% unclustered, cluster := 'unclustered'] 
  
} else {
  
  message('No genes were removed from their corresponding clusters.')
  
}
#
```


### Merge cluster stats
```{r}

all_merged_stats <- merge(merged_stats, sil_df, by=c('gene', 'cluster'), all=T)

```


### Assigning genes to clusters
```{r}
## Factorize
cluster_dt[,gene := factor(gene, levels = ORFs)]

## Add cluster info to counts
orf.perc.mean  <- merge(orf.perc.mean, cluster_dt, by='gene', all.x=T)
orf.sum.perc   <- merge(orf.sum.perc,  cluster_dt, by='gene', all.x=T)

# LAbels frequency
cluster.freq    <- data.table(cluster_dt)[,.N, by=.(cluster) ][order(cluster)]


### Summarise based on literature kinetic class
kin.class.sum   <- orf.sum.perc[,.(sum_read_count=sum(read_count), mean_read_count=mean(read_count), sd_read_count=sd(read_count)), 
                              by=.(Kinetic_class, hpi, Time, cell_line, seqnames, norm_base, sample, group)]


### Summarise based on de-novo kinetic clusters
cluster.sum <- orf.sum.perc[,.(sum_read_count=sum(read_count), mean_read_count=mean(read_count), sd_read_count=sd(read_count)), 
                              by=.(cluster, hpi, Time, cell_line, seqnames, norm_base, sample, group)]

cluster.sum.hpi  <- orf.sum.perc[,.(sum_read_count=sum(read_count), mean_read_count=mean(read_count), sd_read_count=sd(read_count)), 
                              by=.(cluster, hpi, Time, seqnames)]

cluster.sum.hpi.cast <- dcast(cluster.sum.hpi, cluster ~ hpi, value.var = 'mean_read_count')

cluster_dt    <- unique(orf.perc.mean[ , .(gene, Kinetic_class, cluster  )])
class.dt      <- cluster_dt[   , .N, by=.(Kinetic_class, cluster )][order(cluster)]


cluster_dt    <- merge(cluster_dt, all_merged_stats[,.(gene, au, si, gene_to_centroid_dist, sil_width)], by='gene', all.x=T)
```



```{r}
cluster.cds.df <- as.data.table(copy(cluster_dt))
cluster.cds.df[, gene := gsub('_.*', '', gene)]

setDT(CDS.df)
CDS.df[, gene := str_to_upper(gene)]

cluster.cds.df <- merge(CDS.df, 
                        cluster.cds.df, by='gene', all=T)
```


### Summarize and melt
```{r}

orf.sum.perc.cast  <- dcast(orf.sum.perc, gene + Kinetic_class + cluster ~ sample, value.var = 'read_count')


orf.perc.mean.cast <- dcast(orf.perc.mean, gene + Kinetic_class + cluster ~ group, value.var = 'mean')
orf.perc.mean.melt <- melt.data.table(orf.perc.mean.cast, id.vars = c('gene', 'Kinetic_class', 'cluster'), variable.name = 'group', value.name = 'mean')

```




## ORF kinetics, according to literature-based kinetic classes {.tabset}
Time-dependent normalized ORF counts (based on TSSs and TES clusters)


```{r}
### ORFs in samples calculated from TRS-B transcripts with and without leaders

max_ncol <- 10
row_multip <- 10
col_multip <- 4

breaks <- na.omit(unique(metafilt$Time)[order(unique(metafilt$Time))])
labels <- c('', '1', '', '2', '', '4', '6', '8', '12')


plotfun <- function(DT, 
                    geom1 = geom_pointrange(aes(ymin = mean-sd, ymax = mean+sd, color = cell_line)),
                    geom2 = geom_line(aes(x=Time, y=mean, color=cell_line)),
                    #geom2 = geom_smooth(aes(x=Time, y=mean, color=cell_line)),
                    #geom3 = geom_point(aes(x=Time, y = mean, fill=cell_line), shape = 21, size = 3, colour = 'black'),
                    geom3 = NULL,
                    ncol = 10, title = NULL, #"Normalized ORF counts", 
                    strip.backgr = pal_d3()(10)[2],
                    kincolor = strip.backgr,
                    legendpos = 'right',
                    ...
                    ) {
  
  if (nrow(DT) == 0) { 
    ggp <- NULL } else {   # try({ 
      
    # 
    gene_N <- length(unique(DT[,gene]))
    # class_freq <- DT[best_label == 5][,.N,by=class_col]
    
    #colid <- c('gene', class_col)
    #class_freq <- unique(DT[best_label == 5, ..colid])
    
    
    #  min(class_freq)
     # =='adaptive') {}
        
    ggp <- 
      ggplot(DT[
                                #  orf.perc.mean$hpi    != 'dRNA' 
                                #& orf.perc.mean$strand == '+'
                                , ], 
                     aes(Time, mean, color = 'cell_line')
                     #, add = 'loess'
                     ) + 
      geom1 +
      geom2 + 
      geom3 +
      
      scale_color_manual(values = palette) +
      scale_fill_manual(values = palette) +
      scale_y_continuous(labels = scales::percent_format(), name = 'Mean abundance %') +
      scale_x_continuous(breaks = breaks, labels=labels, minor_breaks = NULL, name = 'Hours past infection') +
      
      theme_ipsum() +
      theme(plot.margin = unit(c(1,1,1,1), 'mm')) +
      facet_wrap(~ Kinetic_class + gene, scales = 'free_y', ncol = ncol) +
      theme(plot.margin = unit(c(1,5,1,1), 'mm'),
            legend.position = legendpos,
            legend.text  = element_text(size=13),
            axis.text.x  = element_text(face='bold', size=14),
            axis.text.y  = element_text(face='bold', size=14),
            #axis.title.x = element_text(size=13),
            #axis.title.y = element_text(size=13),
            panel.spacing = unit(2, 'mm'),
            strip.text = element_text(face='bold', size=16),
            strip.background = element_rect(fill = alpha(strip.backgr, 0.4)),
            ...
            )
      # + ggtitle(title)
  
  if (gene_N < ncol) { 
    ggp <- cowplot::plot_grid(ggp, NULL, rel_widths = c(gene_N, ncol - gene_N ), nrow=1)
  }
  
  }   # })
  
  return(ggp)

}
```


```{r}
# Kinetic class frequency

kin_classes      <- unique(orf.perc.mean[,.(gene, Kinetic_class)])
all.class.freq   <- data.table(kin_classes)[,.N, by=.(Kinetic_class) ][order(Kinetic_class)]

my_pal <- pal_d3()(10)

mapfun <- function(i, DT, all.class.freq, palette = my_pal, ...) {
  kin_class <- all.class.freq[i, Kinetic_class]
  strip.background <- palette[i]
  plotfun(DT[Kinetic_class == kin_class],  strip.backgr = strip.background, ...)
}


ncol     <- ifelse(max(all.class.freq$N) < max_ncol, max(all.class.freq$N), max_ncol)
relh     <- ceiling(all.class.freq$N / ncol)  #c(6,1,1,1,2)
figw     <- ncol * col_multip
figh     <- nrow(all.class.freq) * row_multip

```


### Mean and SD with linear connection
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- purrr::map(seq_along(relh), mapfun, orf.perc.mean, all.class.freq, ncol=ncol, 
                  geom1 = geom_pointrange(aes(ymin = mean-sd, ymax = mean+sd, color = cell_line)),
                  geom2 = geom_line(aes(x=Time, y=mean, color=cell_line)),
                  geom3 = NULL)

ggp <- cowplot::plot_grid(plotlist = ggp, ncol = 1, rel_heights = relh)


ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Kinetic.Classes', '_mean.sd.linear.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp
```

### Each replicate with loess
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- purrr::map(seq_along(relh), mapfun, orf.sum.perc, all.class.freq, ncol=ncol, 
                  geom1 = geom_point( aes(x=Time, y=read_count, fill=cell_line), shape = 21, color='black'),
                  geom2 = geom_smooth(aes(x=Time, y=read_count, color=cell_line)),
                  geom3 = NULL)

ggp <- cowplot::plot_grid(plotlist = ggp, ncol = 1, rel_heights = relh)


ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Kinetic.Classes', '_replicates.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp
```


### Mean and SD with loess function
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=F}

ggp <- purrr::map(seq_along(relh), mapfun, orf.perc.mean, all.class.freq, ncol=ncol)

ggp <- cowplot::plot_grid(plotlist = ggp, ncol = 1, rel_heights = relh, align = 'v', axis=c('tlrb'))

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method,  '_Kinetic.Classes', '_mean.sd.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp
```



## ORF kinetics, according to de-novo kinetic classes {.tabset}


```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

denovo_clusters  <- unique(cluster_dt[,.(gene, cluster)])
cluster.freq     <- denovo_clusters[,.N, by=.(cluster) ][order(cluster)]

```



### Figure 5 REVIEW - Mean and SD with linear connection

```{r}
## Scatter and line plots
ncol     <- ifelse(max(cluster.freq$N) < max_ncol, max(cluster.freq$N), max_ncol)
relh     <- ceiling(cluster.freq$N / ncol)  #c(6,1,1,1,2)

plot.axis.title.y <- c(6)
plot.axis.title.x <- c(12)


mapfun <- function(i, DT, label.freq, palette = colorvec, ...) {
  kin_class <- label.freq[i, cluster]
  strip.background <- palette[i]
  
  
  if(i %in% plot.axis.title.y) {
    
    if(i %in% plot.axis.title.x) {
    
      plotfun(DT[cluster   == kin_class],  
              strip.backgr = strip.background, 
              kincolor     = strip.background,
              axis.title.y = element_text(size=13),
              axis.title.x = element_text(size=13),
              ...)
      } else {
        plotfun(DT[cluster   == kin_class],  
              strip.backgr = strip.background, 
              kincolor     = strip.background,
              axis.title.y = element_text(size=13),
              axis.title.x = element_blank(),
              ...)
      }
  } else {
    
    if(i %in% plot.axis.title.x) {
    
      plotfun(DT[cluster   == kin_class],  
              strip.backgr = strip.background, 
              kincolor     = strip.background,
              axis.title.y = element_blank(),
              axis.title.x = element_text(size=13),
              ...)
      } else {
        plotfun(DT[cluster   == kin_class],  
              strip.backgr = strip.background, 
              kincolor     = strip.background,
              axis.title.y = element_blank(),
              axis.title.x = element_blank(),
              ...)
      }
  }
  
}


figw     <- ncol * col_multip
figh     <- 36 ##nrow(cluster.freq) * row_multip
```


```{r}

# plot with only the legend
ggl_legend <- ggplot(cluster.freq, aes(x = 1, y = 1, fill = cluster)) +
  geom_point(shape=21, size=5) +  
  scale_fill_manual(values = palette) +
  theme_void() +  # Remove all other plot components
  theme(
    legend.position = 'none',
    strip.text.x = element_text(size = 10)  
  ) + facet_wrap(~cluster)


```


```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- cowplot::plot_grid(

  cowplot::plot_grid(
    mapfun(1, orf.perc.mean, cluster.freq, palette = palette, legendpos='right', ncol=7),
    nrow = 1, rel_widths = c(1), align = 'vh', axis=c('tblr')
  ),
  
  cowplot::plot_grid(
    mapfun(2, orf.perc.mean, cluster.freq, palette = palette, legendpos='right', ncol=4),
    NULL, 
    mapfun(3, orf.perc.mean, cluster.freq, palette = palette, legendpos='right', ncol=2),
    nrow = 1, rel_widths = c(4, 1, 2), align = 'vh', axis=c('tblr')
  ),
  
  cowplot::plot_grid(
    mapfun(4, orf.perc.mean, cluster.freq, palette = palette, legendpos='right', ncol=7),
    nrow = 1, rel_widths = c(1), align = 'vh', axis=c('tblr')
  ),
  
  cowplot::plot_grid(
    mapfun(5,  orf.perc.mean, cluster.freq, palette = palette, legendpos='right', ncol=7),
    nrow = 1, rel_widths = c(1), align = 'vh', axis=c('tblr')
  ),
  
  cowplot::plot_grid(
    mapfun(6, orf.perc.mean, cluster.freq, palette = palette, legendpos='right', ncol=7),
    nrow = 1, rel_widths = c(1), align = 'vh', axis=c('tblr')
  ),
  
  cowplot::plot_grid(
    mapfun(7, orf.perc.mean, cluster.freq, palette = palette, legendpos='right', ncol=3),
    NULL,
    ggl_legend, ## Heres the color coding ~ aka legend
    nrow = 1, rel_widths = c(3, 3, 1), align = 'vh', axis=c('tblr')
  ),
  
  rel_heights = c(4,1,2,2,2,1), ncol = 1, align = 'vh', axis=c('tblr')

)

ggsave('Figures/Supp Fig S2.jpg',  ggp, width = figw*0.8, height = figh, limitsize = F)

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_DeNovo.Clusters', '_mean.sd.linear.jpg'), ggp, width = figw, height = figh, limitsize = F)


ggp

```


```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=F}
ggp <- purrr::map(seq_along(relh), mapfun, orf.perc.mean, cluster.freq, palette = palette, ncol=ncol)

ggp <- cowplot::plot_grid(plotlist = ggp, ncol = 1, rel_heights = relh, align = 'v', axis=c('tblr'))


ggsave('Figures/Supp Figure S2 TSS Kinetics DeNovo Clusters REVIEW.jpg',  ggp, width = figw*0.8, height = figh, limitsize = F)

ggp
```




```{r}
# 


mapfun <- function(i, DT, cluster.freq, palette = NULL, ...) {
  clustertoplot <- cluster.freq[i, cluster]
  strip.background <- palette[i]
  plotfun(DT[cluster == clustertoplot],  strip.backgr = strip.background, ...)
}


ncol     <- ifelse(max(cluster.freq$N) < max_ncol, max(cluster.freq$N), max_ncol)
relh     <- ceiling(cluster.freq$N / ncol)  #c(6,1,1,1,2)

figw     <- ncol * col_multip
figh     <- nrow(all.class.freq) * row_multip

```


### Mean and SD with linear connection
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- purrr::map(seq_along(relh), mapfun, orf.perc.mean, cluster.freq, palette, ncol=ncol, 
                  geom1 = geom_pointrange(aes(ymin = mean-sd, ymax = mean+sd, color = cell_line)),
                  geom2 = geom_line(aes(x=Time, y=mean, color=cell_line)),
                  geom3 = NULL)

ggp <- cowplot::plot_grid(plotlist = ggp, ncol = 1, rel_heights = relh)


ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_deNovo.Clusters', '_mean.sd.linear.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp
```

### Each replicate with loess
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- purrr::map(seq_along(relh), mapfun, orf.sum.perc, cluster.freq, palette, ncol=ncol, 
                  geom1 = geom_point( aes(x=Time, y=read_count, fill=cell_line), shape = 21, color='black'),
                  geom2 = geom_smooth(aes(x=Time, y=read_count, color=cell_line)),
                  geom3 = NULL)

ggp <- cowplot::plot_grid(plotlist = ggp, ncol = 1, rel_heights = relh)


ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_deNovo.Clusters', '_replicates.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp
```


### Mean and SD with loess function
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=F}

ggp <- purrr::map(seq_along(relh), mapfun, orf.perc.mean, cluster.freq, palette, ncol=ncol)

ggp <- cowplot::plot_grid(plotlist = ggp, ncol = 1, rel_heights = relh, align = 'v', axis=c('tlrb'))

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method,  '_deNovo.Clusters', '_mean.sd.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp
```




## Combine Classes and Clusters
```{r}
figw <- 16
figh <- 8
ncol <- 4
```

sum up all the (normalized) gene counts for each kinetic class in each sample (replicate)
then take the mean and SD of the sums for each hpi and cell line


### Literature Kinetic Classes {.tabset}

```{r}
### Summarise based on literature kinetic class for each sample
kin.class.sum   <- orf.sum.perc[,.(sum_read_count=sum(read_count)),  
                               by=.(seqnames, Kinetic_class, hpi, Time, cell_line, group, sample, norm_base)]

### take the mean for each hpi and cell line (without sample)
kin.class.sum[, `:=`(
  mean_read_count = mean(sum_read_count), sd_read_count   = sd(sum_read_count) ), 
  by = .(seqnames, Kinetic_class, hpi, Time, cell_line)]
                                

```


#### Sum and Loess with Confidence
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- 
      ggplot(kin.class.sum,
             aes(Time, sum_read_count, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_point() +
      geom_smooth() + 
      scale_color_manual(values = palette[]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      scale_y_continuous(labels = scales::percent_format()) +
      scale_x_continuous(breaks = breaks, labels=labels) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      # facet_nested(rows=vars(norm.method), cols=vars(labels), scales = "free_y") ## cols=vars(norm.method)
      facet_wrap(~ Kinetic_class, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.Literature.Classes.', '_points.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)

```

#### Mean and SD with Linear connection
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- 
      ggplot(kin.class.sum,
             aes(Time, mean_read_count, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_line() + 
      geom_pointrange(aes(ymin = mean_read_count - sd_read_count, ymax = mean_read_count + sd_read_count, fill = cell_line), shape = 21, color = 'black' ) +
      scale_color_manual(values = palette[]) +
      scale_fill_manual(values = palette[]) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      scale_y_continuous(labels = scales::percent_format()) +
      scale_x_continuous(breaks = breaks, labels=labels) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      # facet_nested(rows=vars(norm.method), cols=vars(labels), scales = "free_y") ## cols=vars(norm.method)
      facet_wrap(~ Kinetic_class, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.Literature.Classes.', '_means.sd.line.jpg'), ggp, width = figw, height = figh, limitsize = F)

```


### deNovo clusters {.tabset}
```{r}
### Summarise based on literature kinetic class for each sample
denovo.clust.sum  <- orf.sum.perc[,.(sum_read_count=sum(read_count)),  
                               by=.(seqnames, cluster, hpi, Time, cell_line, group, sample, norm_base)]

### take the mean for each hpi and cell line (without sample)
denovo.clust.sum[, `:=`(
  mean_read_count = mean(sum_read_count), sd_read_count   = sd(sum_read_count) ), 
  by = .(seqnames, cluster, hpi, Time, cell_line)]
                                               

```

#### Sum and Loess with Confidence
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- 
      ggplot(denovo.clust.sum,
             aes(Time, sum_read_count, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_point() +
      geom_smooth() + 
      scale_color_manual(values = palette[]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      scale_y_continuous(labels = scales::percent_format()) +
      scale_x_continuous(breaks = breaks, labels=labels) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      # facet_nested(rows=vars(norm.method), cols=vars(labels), scales = "free_y") ## cols=vars(norm.method)
      facet_wrap(~ cluster, scales = 'free_y', ncol = 4)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.DeNovo.Clusters.', '_points.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)

```

#### Mean and SD with Linear connection
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- 
      ggplot(denovo.clust.sum,
             aes(Time, mean_read_count, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_line() + 
      geom_pointrange(aes(ymin = mean_read_count - sd_read_count, ymax = mean_read_count + sd_read_count, fill = cell_line), shape = 21, color = 'black' ) +
      scale_color_manual(values = palette[]) +
      scale_fill_manual(values = palette[]) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      scale_y_continuous(labels = scales::percent_format()) +
      scale_x_continuous(breaks = breaks, labels=labels) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      # facet_nested(rows=vars(norm.method), cols=vars(labels), scales = "free_y") ## cols=vars(norm.method)
      facet_wrap(~ cluster, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.DeNovo.Clusters.', '_means.sd.line.jpg'), ggp, width = figw, height = figh, limitsize = F)

```



## Compare literature and de novo kinetic classes {.tabset}

*NA shows the number of genes where the clustering did not yield an unambiguous result.*
  
### each gene 1
```{r, fig.show='hold',  fig.width = 9, fig.height = 16, eval=T}
ggplot(cluster_dt,
       aes(x=Kinetic_class, y=gene, fill=as.factor(cluster))) +
  geom_tile() +
  # coord_flip() +
  scale_fill_manual(values = palette) +
  #scale_y_discrete(breaks=NULL) +
  theme_ipsum() +
  theme(legend.position = 'bottom',
        panel.grid.major  = element_blank(),
        axis.text.y = element_text(face='italic')) # + facet_wrap(~cluster, ncol = 10)
```

### each gene 2
```{r, fig.show='hold',  fig.width = 9, fig.height = 16, eval=T}
ggplot(cluster_dt,
       aes(x=cluster, y=gene, fill=as.factor(Kinetic_class))) +
  geom_tile() +
  # coord_flip() +
  scale_fill_manual(values = my_pal) +
  #scale_y_discrete(breaks=NULL) +
  theme_ipsum() +
  theme(legend.position = 'bottom',
        panel.grid.major  = element_blank(),
        axis.text.y = element_text(face='italic')) # + facet_wrap(~cluster, ncol = 10)


```

### summary
```{r, fig.show='hold',  fig.width = 18, fig.height = 6, eval=T}
ggplot(class.dt,
       aes(x=Kinetic_class, y=N, fill=as.factor(cluster))) +
  geom_col(position = 'dodge') +
  scale_fill_manual(values = palette) +
  #scale_y_continuous(labels = scales::percent_format()) +
  theme_ipsum() +
  facet_wrap(~cluster, nrow = 1)
```


```{r, fig.show='hold',  fig.width = 18, fig.height = 6, eval=T}
ggplot(class.dt,
       aes(x=as.factor(cluster), y=N, fill=Kinetic_class)) +
  geom_col(position = 'stack') +
  scale_fill_manual(values = my_pal) +
  #scale_y_continuous(labels = scales::percent_format()) +
  theme_ipsum() 
  #+facet_nested_wrap(~cluster, nrow = 1)


```



## Literature Clusters and Selected Genes {.tabset}
```{r}
### Summarise based on literature kinetic class
select.genes.class.sum <- 
  orf.perc.sum[!gene %in% c('US1', 'EP0', 'IE180'),
               .(sum_read_count=sum(read_count)), 
               by=.(Kinetic_class, hpi, Time, cell_line, seqnames, norm_base, sample, group)]

select.genes.class.sum <- rbind(
  select.genes.class.sum,
  orf.perc.sum[gene %in% c('US1', 'EP0', 'IE180'),
               .(sum_read_count=sum(read_count)), 
               by=.(gene, hpi, Time, cell_line, seqnames, norm_base, sample, group)],
  fill = T
)

select.genes.class.sum[gene %in% c('US1', 'EP0', 'IE180'), Kinetic_class := gene]


### take the mean for each hpi and cell line (without sample)
select.genes.class.cell.lines  <- select.genes.class.sum[, .(mean_cell_line=mean(sum_read_count), sd_cell_line=sd(sum_read_count)),
                                by=.(seqnames, Kinetic_class, hpi, Time, cell_line, group)]



#select.genes.class.sum[,mean_cell_line := mean(mean_read_count), by=.(seqnames, Kinetic_class, hpi, Time, cell_line, group)]

#select.genes.class.sum[,sd_cell_line   := sd(mean_read_count), by=.(seqnames, Kinetic_class, hpi, Time, cell_line, group)]


#select.genes.class.cell.lines <- unique(select.genes.class.sum[, .(mean_cell_line, sd_cell_line, seqnames, Kinetic_class, hpi, Time, cell_line, group)] )

```


```{r}
figw <- 16
figh <- 10
ncol <- 10
```


### Reps with loess and confidence
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ncol=3

ggp <- 
      ggplot(select.genes.class.sum,
             aes(Time, sum_read_count, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_point() +
      geom_smooth() + 
      scale_color_manual(values = palette[]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      scale_y_continuous(labels = scales::percent_format()) +
      scale_x_continuous(breaks = breaks, labels=labels) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      #facet_nested(cols=vars(Kinetic_class), scales = "free_y") ## cols=vars(norm.method), rows=vars(norm.method), 
      facet_wrap(~ Kinetic_class, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.Literature.Clusters.SELECT', '_points.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)
```

### Means with SD and linear connection
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggp <- 
      ggplot(select.genes.class.cell.lines,
             aes(Time, mean_cell_line, color = cell_line)
                     #, add = 'loess'
                     ) + 
      geom_pointrange(aes(Time, ymin=mean_cell_line - sd_cell_line, ymax=mean_cell_line + sd_cell_line, color = cell_line)) +    
      geom_point() +
      geom_line(aes(group = cell_line)) + 
      scale_color_manual(values = palette[]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      scale_y_continuous(labels = scales::percent_format()) +
      scale_x_continuous(breaks = breaks, labels=labels) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      #facet_nested(cols=vars(Kinetic_class), scales = "free_y") ## cols=vars(norm.method), rows=vars(norm.method),
      facet_wrap(~ Kinetic_class, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.Literature.Clusters.SELECT', '_mean.SD.linear.jpg'), ggp, width = figw, height = figh, limitsize = F)
```



## Gene-pair ratios {.tabset}

These pairs were analyzed: 
ie180/us1, ie180/ep0, us1/ep0, noir1/noir2, noir1/ie180, noir2/ie180

```{r, fig.show='hold',  fig.width = 10, fig.height = 7, eval=T}
ab.US1    <- setnames(orf.perc.sum[gene == 'US1',    .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'US1')

ab.IE180  <- setnames(orf.perc.sum[gene == 'IE180',  .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'IE180')

ab.EP0    <- setnames(orf.perc.sum[gene == 'EP0',    .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'EP0')

ab.noir   <- setnames(orf.perc.sum[gene == 'NOIR',   .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'NOIR')

ab.noir2  <- setnames(orf.perc.sum[gene == 'NOIR-2', .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'NOIR-2')


ab.merged <- merge(ab.US1,    ab.IE180,  by=c('sample', 'hpi', 'Time', 'cell_line', 'group', 'norm.method'))
ab.merged <- merge(ab.merged, ab.EP0,    by=c('sample', 'hpi', 'Time', 'cell_line', 'group', 'norm.method'))
ab.merged <- merge(ab.merged, ab.noir,   by=c('sample', 'hpi', 'Time', 'cell_line', 'group', 'norm.method'))
#ab.merged <- merge(ab.merged, ab.noir2,  by=c('sample', 'hpi', 'Time', 'cell_line', 'norm.method'))

ab.merged[, IE180_US1   :=  IE180   /  US1]
ab.merged[, IE180_EP0   :=  IE180   /  EP0]
ab.merged[, US1_EP0     :=  US1     /  EP0]
#ab.merged[, NOIR1_NOIR2 :=  NOIR    / `NOIR-2`]
ab.merged[, NOIR_IE180  :=  NOIR    /  IE180]
#ab.merged[, NOIR2_IE180 := `NOIR-2` /  IE180]

#ab.merged[, US1_IE180   :=  US1     / IE180     ]
#ab.merged[, EP0_IE180   :=  EP0     / IE180     ]
#ab.merged[, EP0_US1     :=  EP0     / US1       ]
#ab.merged[, NOIR2_NOIR1 := `NOIR-2` / NOIR      ]
ab.merged[, IE180_NOIR  :=  IE180   / NOIR      ]
#ab.merged[, IE180_NOIR2 :=  IE180   / `NOIR-2`  ]

gpairs <- c(
'IE180_US1' , 
'IE180_EP0'  , 
'US1_EP0'     ,
'NOIR1_NOIR2' ,
'NOIR_IE180'  ,
'NOIR2_IE180' ,

'US1_IE180'   ,
'EP0_IE180'   ,
'EP0_US1'     ,
'NOIR2_NOIR1' ,
'IE180_NOIR'  ,
'IE180_NOIR2' 
)
gpairs <- gsub('_', '/', gpairs)

ab.merged.gt <- ab.merged#[,-c(7:11)]

ab.merged.gt <- melt(ab.merged.gt, variable.name = 'gene_pair', value.name = 'ratio', id.vars = c(1:6))

ab.merged.gt$gene_pair <- gsub('_', '/', ab.merged.gt$gene_pair)

ab.merged.gt <- ab.merged.gt[gene_pair %in% gpairs, ]

ab.merged.gt[,gene_pair := factor(gene_pair, levels = gpairs)]

ab.merged.gt[,mean := mean(ratio),   by=.(hpi, Time, cell_line, group, norm.method, gene_pair)]
ab.merged.gt[,sd   := sd  (ratio),   by=.(hpi, Time, cell_line, group, norm.method, gene_pair)]
ab.merged.gt[,sum  := sum (ratio),   by=.(hpi, Time, cell_line, group, norm.method, gene_pair)]



plot_data <- ab.merged.gt[norm.method == 'viral_read.count'
                 # norm.method == 'host_read.count' 
                 , ]
```


```{r}
figw <- 16
figh <- 10
ncol <- 10
```


### Means with SD and linear connection
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}
ggmean <- ggplot(plot_data) + 
  #geom_point(aes(Time, ratio, color=cell_line)) +
  geom_line(aes(Time, mean, color=cell_line, group=cell_line)) +
  geom_pointrange(aes(Time, mean, color=cell_line, group=cell_line, ymin=mean-sd, ymax=mean+sd)) +
  scale_color_manual(values = palette[]) +
  scale_y_continuous(name='ratio') +
  scale_x_continuous(breaks = breaks, labels=labels) +
  theme_ipsum() +
  #facet_nested(rows = vars(norm.method), cols=vars(Time))
  facet_wrap(~gene_pair, scales = 'free', ncol=3)


ggmean
ggsave(paste0(fig.dir, '/gene_ratios_mean.jpg'), width = 10, height = 7)
```



### Reps with loess and confidence
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

ggrep <- ggplot(plot_data[# norm.method == 'viral_read.count'
                 # norm.method == 'host_read.count' 
                 , ]) + 
  geom_point(aes(Time, ratio, color=cell_line)) +
  geom_smooth(aes(Time, ratio, color=cell_line, group=cell_line)) +
  #geom_pointrange(aes(Time, mean, color=cell_line, group=cell_line, ymin=mean-sd, ymax=mean+sd)) +
  scale_color_manual(values = palette[]) +
  scale_y_continuous(name='ratio') +
  scale_x_continuous(breaks = breaks, labels=labels) +
  theme_ipsum() +
  #facet_nested(rows = vars(norm.method), cols=vars(Time))
  facet_wrap(~gene_pair, scales = 'free', ncol=3)


ggrep
ggsave(paste0(fig.dir, '/gene_ratios_rep.jpg'), width = 10, height = 7)
```


### Sums with linear connection
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}
ggsum <- ggplot(plot_data) + 
  geom_point(aes(Time, sum, color=cell_line)) +
  geom_line(aes(Time, sum, color=cell_line, group=cell_line)) +
  #geom_pointrange(aes(Time, mean, color=cell_line, group=cell_line, ymin=mean-sd, ymax=mean+sd)) +
  scale_color_manual(values = palette[]) +
  scale_y_continuous(name='ratio') +
  scale_x_continuous(breaks = breaks, labels=labels) +
  theme_ipsum() +
  #facet_nested(rows = vars(norm.method), cols=vars(Time))
  facet_wrap(~gene_pair, scales = 'free', ncol=3)


ggsum
ggsave(paste0(fig.dir, '/gene_ratios_sum.jpg'), width = 10, height = 7)

```


```{r}
## Export data
if (writetables) {
  fwrite(orf.sum.perc,          paste0(fig.dir, '/', norm.method, '_', EndType, '_abund.tsv'), sep = '\t')
  fwrite(orf.sum.perc.cast,     paste0(fig.dir, '/', norm.method, '_', EndType, '_abund.cast.tsv'), sep = '\t')
  fwrite(orf.perc.mean,         paste0(fig.dir, '/', norm.method, '_', EndType, '_abund.mean.tsv'), sep = '\t')
  fwrite(orf.perc.mean.cast,    paste0(fig.dir, '/', norm.method, '_', EndType, '_abund.mean.cast.tsv'), sep = '\t')
  
  fwrite(all_merged_stats,      paste0(fig.dir, '/', norm.method, '_', EndType, '_all_cluster_stats.tsv'), sep = '\t')
  fwrite(cluster.cds.df,        paste0(fig.dir, '/', norm.method, '_', EndType, '_cluster.cds.df.tsv'), sep = '\t')
  fwrite(cluster_dt,            paste0(fig.dir, '/', norm.method, '_', EndType, '_Hclust.genes.and.kin.classes.tsv'), sep = '\t')
  fwrite(cluster.sum.hpi.cast,  paste0(fig.dir, '/', norm.method, '_', EndType, '_cluster.sum.hpi.cast.tsv'), sep = '\t')
}
```


```{r bib, include=FALSE}
# KEEP THIS AT THE END OF THE DOCUMENT TO GENERATE A LOCAL bib FILE FOR PKGS USED
knitr::write_bib(sub("^package:", "", grep("package", search(), value=TRUE)), file='skeleton.bib')
```