---
title: "PRV - 4 cell lines: TSS dynamics"
author: Balázs Kakuk
output: 
  html_document:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: false
---

```{css, echo=FALSE}
    body .main-container {
      max-width: 90% !important;
      width: 90% !important;
    }
    body {
      max-width: 90% !important;
      margin-left: auto;
      margin-right: auto;
    }
```


```{r include=FALSE}
## hrbrthemes::ipsum:

knitr::opts_chunk$set(fig.retina=2, echo = FALSE, fig.align = 'center', message = T, warning = F)
```

```{r ipsum_setup, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE, include=FALSE}
library(prettydoc)
library(hrbrthemes, quietly = T)
#library(GenomicFeatures, quietly = T)
library(DESeq2, quietly = T)
library(ggpubr, quietly = T)
#library(misc, quietly = T)
library(dplyr, quietly = T)
#library(tidyverse, quietly = T)
library(ggsci, quietly = T)
#library(Gviz, quietly = T)
library(dplyr, quietly = T)
library(tidyr)
library(scales)
#library(plyr)
#library(moanin, quietly = T)
library(grid, quietly = T)
library(gridExtra, quietly = T) 
library(ggpubr, quietly = T)
library(data.table)
library(ggh4x)
library(knitr)
library(formattable)



### Project name and output directory
outdir  <- 'PRV-MDBIO-4cell'; try({ dir.create(outdir) })

#### Load configuration file
config <- readRDS(paste0(outdir, '_config.rds'))

####
source('_WF.part0.R')

#### Reference Transcripts

### Import and Format 
source('import.ref.TRs.R')

### Export mRNAS and genes (CDSs)
export.gff2(as.data.frame(viral.ref), "refgenome/LT934125.1_Torma_et_al_corrected_w_CDS.genes.gff2")
export.gff3(as.data.frame(viral.ref), "refgenome/LT934125.1_Torma_et_al_corrected_w_CDS.genes.gff3")

fwrite(TR.Ref.data, paste0(outdir, '/TR.Ref.data.tsv'), sep = '\t')
TR.Ref.data <- fread(paste0(outdir, '/TR.Ref.data.tsv'))



#### CAGE
outdir  <- 'CAGE'; try({ dir.create(outdir) })
  
##config
source('CAGE_config.R')
  
####
outdir  <- 'PRV-MDBIO-4cell'; try({ dir.create(outdir) })



#### Settings ####

## Main
#save.data <- 'PRV.rebase.RData'
EndType <- 'TSS'

#project_config  <- fread('project_config.txt')
#outdir  <- project_config$outdir
res.dir <- outdir; try({ dir.create(res.dir) })
fig.dir <- paste0(res.dir, '/', EndType, '_abund'); # try({ dir.create(fig.dir) })## save plots to this directory

## Miscallenaous
## Colors
colorvec <- na.omit(c(
  colorvec[c(1:5,15,13,10,24,7,26,32,33)],
  colorvec[-c(1:5,15,13,10,24,7,26,32,33)]))

colorvec <- colorvec[-c(9,13)]


colorvec <- colorvec[-c(10,13,17,18)]

colorvec <- colorvec[c(1:2,14,3:length(colorvec))]

palette <- colorvec  #pal_npg()(10)

##
writetables <- T
load.prev   <- F



## Filter metadata for those samples that had at least 1 viral read
readcounts <- fread(paste0(outdir, '/readcounts.tsv'), header = T)
metafilt   <- metafilt[metafilt$sample %in% readcounts$sample, ]


#### Normalization method
norm.method <- 'viral_read.count'
# Normalization before or after filtering for counts that didn't overlap with the other end?
norm_before_filtering <- F

```


# Project info
Samples: dcDNA Seq  
pychopper: no  
mapping version: v6  
LoRTIA: yes (stranded only output)  


```{r}
#
CAGE_TSSr_clusters  <- fread(paste0(CAGE_config$outdir, '/TSSr.CAGE.all.Cluster.Results.txt'))
#
dcDNA_TSSr_clusters <- fread(paste0(config$outdir,      '/TSSr.dcDNA.all.Cluster.Results.txt'))


outdir <- config$outdir


#### dcDNA TSSr
TR.Ref.data.dcDNA        <- fread(paste0(outdir, '/TR.Ref.data.dcDNA.tsv'),        na.strings = '')
TR.gff.compare.uni.dcDNA <- fread(paste0(outdir, '/TR.gff.compare.uni.dcDNA.tsv'), na.strings = '')
dcDNA.support.freq       <- fread(paste0(outdir, '/dcDNA.support.freq.tsv'),       na.strings = '')
dcDNA_TSSr_clusters_uni  <- fread(paste0(outdir, '/dcDNA.TSSr_clusters_uni.tsv'),  na.strings = '')

dcDNA_ClusterShapes      <- fread(paste0(outdir, '/TSSr.dcDNA.all.ClusterShapes.txt'))
##
dcDNA_assignedClusters   <- fread(paste0(outdir, '/TSSr.dcDNA.best.assignedClusters.txt'), na.strings = '')


#### CAGE TSSr
TR.Ref.data.CAGE        <- fread(paste0(outdir, '/TR.Ref.data.CAGE.tsv'),        na.strings = '')
TR.gff.compare.uni.CAGE <- fread(paste0(outdir, '/TR.gff.compare.uni.CAGE.tsv'), na.strings = '')
CAGE.support.freq       <- fread(paste0(outdir, '/CAGE.support.freq.tsv'),       na.strings = '')
CAGE_TSSr_clusters_uni  <- fread(paste0(outdir, '/CAGE.TSSr_clusters_uni.tsv'),  na.strings = '')

####
EX.uni    <- fread(paste0(outdir, '/EX.uni.tsv'),  na.strings = '')

#### dcDNA TransFrags Results
TR.gff.compare.merged.TR.counts.gt <- fread(paste0(outdir, "/TR.gff.compare.merged.TR.counts.gt.tsv"), na.strings = '')

cols_to_group <- setdiff(colnames(TR.gff.compare.merged.TR.counts.gt), c(metacols, "count", 'rep', 'total', 'contig_size', "average_coverage", "norm_count"))
TR.gff.compare.uni <- unique(TR.gff.compare.merged.TR.counts.gt[,..cols_to_group])
dup(TR.gff.compare.uni$transcript_id)

TR.gff.compare.uni[,prime5 := fifelse(strand == '+', start, end)]
TR.gff.compare.uni[,prime3 := fifelse(strand == '+', end, start)]

TR.gff.compare.uni[,TR.prime5.win.start := prime5]
TR.gff.compare.uni[,TR.prime5.win.end   := prime5]
TR.gff.compare.uni[,TR.prime3.win.start := prime3]
TR.gff.compare.uni[,TR.prime3.win.end   := prime3]


```

# TSSr Clustering results of CAGE and dcDNA data

```{r, fig.width=32, fig.height=6, eval=F}
dcDNA_TSSr_clusters_uni[,TC.width := (TC.end - TC.start) + 1]


## histogram
ggplot(dcDNA_TSSr_clusters_uni) +
  geom_rect(#data = dcDNA_TSSr_clusters_uni[strand == '+'],
            aes(xmin = TC.start, xmax = TC.end, ymin = 1, ymax = 2)) +
  #xlim(c(0,250))  +
  facet_nested(rows=vars(strand)) +
  theme_minimal() +
  theme(legend.position = 'bottom')


```

## Cluster Widths

### dcDNA
```{r}
dcDNA_TSSr_clusters_uni[,TC.width := (TC.end - TC.start) + 1]


## histogram
ggplot(dcDNA_TSSr_clusters_uni) +
  geom_histogram(aes(TC.width), binwidth =  10) +
  #xlim(c(0,250))  +
  #facet_nested(rows=vars(hpi)) +
  theme_minimal() +
  theme(legend.position = 'bottom')


```

### CAGE
```{r}
CAGE_TSSr_clusters_uni[,TC.width := (TC.end - TC.start) + 1]


## histogram
ggplot(CAGE_TSSr_clusters_uni) +
  geom_histogram(aes(TC.width), binwidth =  10) +
  #xlim(c(0,250))  +
  #facet_nested(rows=vars(hpi)) +
  theme_minimal() +
  theme(legend.position = 'bottom')


```

## Assigning clutsers to genes

```{r}
```


### Plot top20 Gene-Wise Shifts
```{r, fig.width=9, fig.height=15}
gene_TSS_shifts_Sum <- unique(dcDNA_assignedClusters[,
                                .(chr, strand, gene, gene_mean_shift_TSS, gene_sd_shift_TSS, gene_varcoeff_shif_TSS)])
gene_TSS_shifts_Sum <- gene_TSS_shifts_Sum[order(gene_mean_shift_TSS, decreasing = T)]
topShifts   <- gene_TSS_shifts_Sum[1:20, gene]
topShifts   <- topShifts[!topShifts %in% c('US1_2', 'IE180_2')]

### Plot Gene-Wise Shifts
ggplot(dcDNA_assignedClusters[gene %in% topShifts
                      & Time %in% c(1,2,4,6,8,12)  ], 
       aes(Time, gene_dominant_TSS_shift)) + 
  geom_line(aes(color = cell_line, group = cell_line)) + 
  geom_point(aes(color = cell_line)) + 
  #coord_flip() +
  facet_nested_wrap(~gene, ncol=3) +
  theme_minimal()



```



# Merge dcDNA TSSr Clustering results and CAGE TSSr clustering results with TransFrags and Reference Transcripts


## The Transcripts were merged based on their TSS with both the dcDNA-based TSS lusters and the CAGE TSS clusters


## Reference Transcripts
```{r}
#TR.Ref.data.TSSr.comb <- copy(TR.Ref.data.CAGE)
cnames <- c('TC.start', 'TC.end', 'support', 'score', 'consensus_peak')
setnames(TR.Ref.data.CAGE,  cnames, paste0('CAGE_',  cnames))
setnames(TR.Ref.data.dcDNA, cnames, paste0('dcDNA_', cnames))

TR.Ref.data.TSSr.comb <- merge(TR.Ref.data.CAGE, TR.Ref.data.dcDNA,
                               by=colnames(TR.Ref.data.CAGE)[2:17])

ggVennDiagram::ggVennDiagram(list(
  REF_TR = TR.Ref.data$transcript_id,
  CAGE_TC_supported   = TR.Ref.data.CAGE[!is.na(CAGE_TC_ID), transcript_id],
  dcDNA_TC_supported  = TR.Ref.data.dcDNA[!is.na(dcDNA_TC_ID), transcript_id]
  ))
```


## TransFrags
```{r}
### TransFrags
#TR.Ref.data.TSSr.comb <- copy(TR.Ref.data.CAGE)
cnames <- c('TC.start', 'TC.end', 'support', 'score', 'consensus_peak')
setnames(TR.gff.compare.uni.CAGE,  cnames, paste0('CAGE_',  cnames))
setnames(TR.gff.compare.uni.dcDNA, cnames, paste0('dcDNA_', cnames))

TransFrags.TSSr.comb <- merge(TR.gff.compare.uni.CAGE, TR.gff.compare.uni.dcDNA,
                               by=colnames(TR.gff.compare.uni.CAGE)[c(2:4,10:27)])


ggVennDiagram::ggVennDiagram(list(
  TransFrag = TR.gff.compare.uni$transcript_id,
  CAGE_TC_supported   = TR.gff.compare.uni.CAGE[!is.na(CAGE_TC_ID),   transcript_id],
  dcDNA_TC_supported  = TR.gff.compare.uni.dcDNA[!is.na(dcDNA_TC_ID), transcript_id]
))
```


```{r}
### Non-spliced Transfrags ONly !!!
TR_EX           <- dcast(EX.uni, seqnames + strand + TR_ID ~ exon_rank, value.var = 'exon_ID')
non.spliced.TFs <- TR_EX[is.na(`2`), TR_ID]

TransFrags.TSSr.comb <- TransFrags.TSSr.comb[transcript_id %in% non.spliced.TFs, ]
```


## Validate TESs of TransFrags
```{r}
### Validate TESs
# make start and end as prime3 for joining
TransFrags.TSSr.comb[,start := prime3]
TransFrags.TSSr.comb[,end   := prime3]

# "Valid" prime3 sites form reference transcripts
#valid.prime3[,TES := start + 10]
#fwrite(valid.prime3, 'valid.prime3.tsv', sep = '\t')

# had to correct, as some TESs were very close ! -->> cehck !!!
valid.prime3 <- fread('valid.prime3.tsv')

prime3.TF.ov <- foverlaps2(TransFrags.TSSr.comb, valid.prime3, by.x=c('seqnames', 'strand', 'start', 'end'), by.y=c('seqnames', 'strand', 'start', 'end'), minoverlap = 1)
prime3.TF.ov <- unique(prime3.TF.ov[,.(seqnames, strand, transcript_id, valid_tes, TES)])

# Merge with TransFrags
TransFrags.TSSr.comb <- merge(TransFrags.TSSr.comb, prime3.TF.ov, by=c('seqnames', 'strand', 'transcript_id'), all.x=T)
TransFrags.TSSr.comb[is.na(valid_tes), valid_tes := F]

# correct start and end
TransFrags.TSSr.comb[, start  := fifelse(strand == '+', prime5, prime3)]
TransFrags.TSSr.comb[, end    := fifelse(strand == '+', prime3, prime5)]

prime3.valid.corr.freq <- TransFrags.TSSr.comb[,.N, by=.(valid_tes)]

ggvf <- ggplot(prime3.valid.corr.freq) +
  geom_col(aes(x=valid_tes, y=N, fill=valid_tes), color='black') +
  theme_bw() +
  ggtitle('TransFrags 3-prime end result, according to ref mRNA TES')

ggsave(file.path(outdir, 'TF_TES_validation.jpg'), ggvf, height = 12, width = 9)


ggvf
```


## Annotate new Transcripts from TSS-validated TransFrags {.tabset}


### TransFrags that are not equal to refernce
```{r}
ggVennDiagram::ggVennDiagram(list(
  TransFrag = TransFrags.TSSr.comb$transcript_id,
  CAGE_TC_supported   = TransFrags.TSSr.comb[!is.na(CAGE_TC_ID)   & class_code != '=',  transcript_id],
  dcDNA_TC_supported  = TransFrags.TSSr.comb[!is.na(dcDNA_TC_ID)  & class_code != '=',  transcript_id]
))
```

### TransFrags that are have at least '**' 
```{r}
ggVennDiagram::ggVennDiagram(list(
  TransFrag = TransFrags.TSSr.comb$transcript_id,
  CAGE_TC_supported   = TransFrags.TSSr.comb[!is.na(CAGE_TC_ID)   & class_code != '=' & CAGE_TC_significance  != '*',  transcript_id],
  dcDNA_TC_supported  = TransFrags.TSSr.comb[!is.na(dcDNA_TC_ID)  & class_code != '=' & dcDNA_TC_significance != '*',  transcript_id]
))
```

### TransFrags that share 3-prime with previous annotation -->> significant drop
```{r}
ggVennDiagram::ggVennDiagram(list(
  TransFrag = TransFrags.TSSr.comb$transcript_id,
  CAGE_TC_supported   = TransFrags.TSSr.comb[!is.na(CAGE_TC_ID)   & class_code != '=' & CAGE_TC_significance  != '*' & valid_tes == T,  transcript_id],
  dcDNA_TC_supported  = TransFrags.TSSr.comb[!is.na(dcDNA_TC_ID)  & class_code != '=' & dcDNA_TC_significance != '*' & valid_tes == T,  transcript_id]
))
```


```{r}
TransFrags.TSSr.filt <-  TransFrags.TSSr.comb[  (!is.na(CAGE_TC_ID)  & CAGE_TC_significance   %in% c('**', '***'))
                                              | (!is.na(dcDNA_TC_ID) & dcDNA_TC_significance  %in% c('**', '***'))
                                              #& class_code != '='  & valid_tes == T
                                              ,  ]

TransFrags.TSSr.filt[,.N, by=.(CAGE_TC_significance, dcDNA_TC_significance)]

TransFrags.TSSr.filt <-  TransFrags.TSSr.filt[class_code != '='  & valid_tes == T, ]

TransFrags.TSSr.filt[,.N, by=.(CAGE_TC_significance, dcDNA_TC_significance, valid_tes)]
```


## Merge TransFrags with TSSr group-wise cluster results (for the peaks)
```{r}
### Merge TransFrags with TSSr group-wise cluster results (for the peaks)
TransFrags.TSSr.filt.counts.gt <- merge(TR.gff.compare.merged.TR.counts.gt[,.(seqnames, strand, transcript_id, sample, rep, group, cell_line, hpi, Time, count, norm_count)], 
                                        TransFrags.TSSr.filt,
                                        by= c('seqnames', 'strand', 'transcript_id'),
                                        all.y = T)

TransFrags.TSSr.filt.counts.gt[, cell_line := gsub('-', '_', cell_line)]

## Merge with dcDNA clusters
TransFrags.TSSr.filt.counts.gt  <- merge(TransFrags.TSSr.filt.counts.gt, 
                                         dcDNA_TSSr_clusters[,.(cluster, seqnames = chr, strand, cell_line, hpi, 
                                                                cluster.start = start, cluster.end = end, dcDNA_TC_ID = paste0('TC_', cluster),
                                                                dominant_tss, tags, tags.dominant_tss,  q_0.1, q_0.9, interquantile_width, shape.score, cluster_width)],
                                         #by.y = c('seqnames', 'strand', 'cluster', 'group'),
                                         by = c('seqnames', 'strand', 'dcDNA_TC_ID', 'cell_line', 'hpi'),
                                         all.x = T )

# Caluclate distance of TransFrags to peak of TC                      
TransFrags.TSSr.filt.counts.gt[, TSS_dcDNA_distance := dominant_tss - prime5, by= transcript_id]

# Reanme columns
cnames <- c('cluster.start', 'cluster.end', 'dominant_tss', 'tags', 'tags.dominant_tss',  'q_0.1', 'q_0.9', 'interquantile_width', 'shape.score', 'cluster_width')
setnames(TransFrags.TSSr.filt.counts.gt,  cnames, paste0('dcDNA_', cnames))


## Merge with CAGE clusters
TransFrags.TSSr.filt.counts.gt  <- merge(TransFrags.TSSr.filt.counts.gt, 
                                         CAGE_TSSr_clusters[,.(cluster, seqnames = chr, strand, cell_line, hpi, 
                                                                cluster.start = start, cluster.end = end, CAGE_TC_ID = paste0('TC_', cluster),
                                                                dominant_tss, tags, tags.dominant_tss,  q_0.1, q_0.9, interquantile_width, shape.score, cluster_width)],
                                         #by.y = c('seqnames', 'strand', 'cluster', 'group'),
                                         by = c('seqnames', 'strand', 'CAGE_TC_ID', 'cell_line', 'hpi'),
                                         all.x = T )

# Caluclate distance of TransFrags to peak of TC                      
TransFrags.TSSr.filt.counts.gt[, TSS_CAGE_distance := dominant_tss - prime5, by= transcript_id]

# Reanme columns
setnames(TransFrags.TSSr.filt.counts.gt,  cnames, paste0('CAGE_', cnames))
```

### Distance to dcDNA cluster's dominant TSS (peak)
```{r}
# Check distances
TransFrags.TSSr.filt.counts.gt[, .N, TSS_dcDNA_distance]
ggplot(TransFrags.TSSr.filt.counts.gt, aes(TSS_dcDNA_distance) ) + geom_histogram(bins=100) + 
  geom_vline(xintercept = c(-10, 10), color='red') +
  geom_vline(xintercept = c(-25, 25))
```

### Distance to CAGE cluster's dominant TSS (peak)
```{r}
TransFrags.TSSr.filt.counts.gt[, .N, TSS_CAGE_distance]
ggplot(TransFrags.TSSr.filt.counts.gt, aes(TSS_CAGE_distance) ) + geom_histogram(bins=100) + 
  geom_vline(xintercept = c(-10, 10), color='red') +
  geom_vline(xintercept = c(-25, 25))
```



## TransFrags that share 5-primes with CAGE or dcDNA TSS cluster peak in the respective sample group with a +/- 10 bp wobble
```{r}
## TransFrags that share 5-primes with CAGE or dcDNA TSS cluster peak in the respective sample group with a +/- 10 bp wobble
ggVennDiagram::ggVennDiagram(list(
  TransFrag           = unique(TransFrags.TSSr.filt.counts.gt$transcript_id),
  CAGE_TC_supported   = unique(TransFrags.TSSr.filt.counts.gt[abs(TSS_CAGE_distance) <= 10,   transcript_id]),
  dcDNA_TC_supported  = unique(TransFrags.TSSr.filt.counts.gt[abs(TSS_dcDNA_distance) <= 10,  transcript_id])
))


TransFrags.TSSr.filt.counts.gt <- TransFrags.TSSr.filt.counts.gt[abs(TSS_CAGE_distance) <= 10 | abs(TSS_dcDNA_distance) <= 10, ]
TransFrags.TSSr.filt.counts.gt[is.na(TSS_CAGE_distance) & is.na(TSS_dcDNA_distance)]
```

*The TSS here will thus be the peak of the TC (either in the CAGE ir dcDNA - which is closer),*  
*considering the +/- 10 bp wobble*

*Filtering for read count : >= 5*  
*Filtering for group prevalence : >= 2*  

```{r}
# The TSS here will thus be the peak of the TC (either in the CAGE ir dcDNA - which is closer)

TransFrags.TSSr.filt.counts.gt[, TSS := fifelse(is.na(TSS_dcDNA_distance), CAGE_dominant_tss, 
                                          fifelse(is.na(TSS_CAGE_distance),  dcDNA_dominant_tss,
                                            fifelse(abs(TSS_dcDNA_distance) < abs(TSS_CAGE_distance), dcDNA_dominant_tss, CAGE_dominant_tss
                                                    ))), by=transcript_id]

TransFrags.TSSr.filt.counts.gt[,.N, by=.(TSS, TSS_dcDNA_distance, TSS_CAGE_distance)]



TransFrags.TSSr.filt.counts.gt <- unique(TransFrags.TSSr.filt.counts.gt[,.(seqnames, strand, start, end, TSS, TES,
                                                                transcript_id, cmp_ref, class_code, 
                                                                group, hpi, Time, cell_line, group, sample, count)])

TransFrags.TSSr.filt.group.counts <- TransFrags.TSSr.filt.counts.gt[,.(sum_count = sum(count)), by=.(seqnames, strand, TSS, TES, group, cell_line, hpi, Time)]

## Add Novel TX ID
TransFrags.TSSr.filt.group.counts[, NOV_TX_ID := paste0('NOV_TX_', .GRP), by = .(seqnames, strand, TSS, TES)]

## Filter for read_count
TransFrags.TSSr.filt.group.counts <- TransFrags.TSSr.filt.group.counts[ sum_count >= 5, ]

## Filter for prevalence?
TransFrags.TSSr.filt.group.counts[, prevalence := uniqueN(group), by = NOV_TX_ID]
TransFrags.TSSr.filt.group.counts <- TransFrags.TSSr.filt.group.counts[ prevalence  >= 2, ]


## 
NOV.TXs.uni <- TransFrags.TSSr.filt.group.counts[,.(sum_count = sum(sum_count)), by = .(seqnames, strand, TSS, TES, NOV_TX_ID) ]

NOV.TXs.uni.sp <- dcast(TransFrags.TSSr.filt.group.counts, seqnames + strand + TSS + TES + NOV_TX_ID ~ group, value.var = 'sum_count')

```


```{r}
## TransFrags that share 5-primes with CAGE or dcDNA TSS cluster peak in the respective sample group with a +/- 10 bp wobble
ggVennDiagram::ggVennDiagram(list(
  PK_15           = unique(TransFrags.TSSr.filt.group.counts[ cell_line == 'PK_15', NOV_TX_ID]),
  PC_12           = unique(TransFrags.TSSr.filt.group.counts[ cell_line == 'PC_12', NOV_TX_ID]),
  C6              = unique(TransFrags.TSSr.filt.group.counts[ cell_line == 'C6',    NOV_TX_ID])
))

```











```{r bib, include=FALSE}
# KEEP THIS AT THE END OF THE DOCUMENT TO GENERATE A LOCAL bib FILE FOR PKGS USED
knitr::write_bib(sub("^package:", "", grep("package", search(), value=TRUE)), file='skeleton.bib')
```