---
title: "PRV - 4 cell lines: TSS dynamics"
author: Balázs Kakuk
output: 
  html_document:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: false
---

```{css, echo=FALSE}
    body .main-container {
      max-width: 90% !important;
      width: 90% !important;
    }
    body {
      max-width: 90% !important;
      margin-left: auto;
      margin-right: auto;
    }
```


```{r include=FALSE}
## hrbrthemes::ipsum:

knitr::opts_chunk$set(fig.retina=2, echo = FALSE, fig.align = 'center', message = T, warning = F)
```

```{r ipsum_setup, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE, include=FALSE}
library(prettydoc)
library(hrbrthemes, quietly = T)
#library(GenomicFeatures, quietly = T)
library(DESeq2, quietly = T)
library(ggpubr, quietly = T)
#library(misc, quietly = T)
library(dplyr, quietly = T)
#library(tidyverse, quietly = T)
library(ggsci, quietly = T)
#library(Gviz, quietly = T)
library(dplyr, quietly = T)
library(tidyr)
library(scales)
#library(plyr)
#library(moanin, quietly = T)
library(grid, quietly = T)
library(gridExtra, quietly = T) 
library(ggpubr, quietly = T)
library(data.table)
library(ggh4x)
library(knitr)
library(formattable)



### Project name and output directory
outdir  <- 'PRV-MDBIO-4cell'; try({ dir.create(outdir) })

#### Load configuration file
config <- readRDS(paste0(outdir, '_config.rds'))

####
source('_WF.part0.R')

#### Reference Transcripts

### Import and Format 
source('import.ref.TRs.R')

### Export mRNAS and genes (CDSs)
export.gff2(as.data.frame(viral.ref), "refgenome/LT934125.1_Torma_et_al_corrected_w_CDS.genes.gff2")
export.gff3(as.data.frame(viral.ref), "refgenome/LT934125.1_Torma_et_al_corrected_w_CDS.genes.gff3")

fwrite(TR.Ref.data, paste0(outdir, '/TR.Ref.data.tsv'), sep = '\t')
TR.Ref.data <- fread(paste0(outdir, '/TR.Ref.data.tsv'))



#### CAGE
outdir  <- 'CAGE'; try({ dir.create(outdir) })
  
##config
source('CAGE_config.R')
  
####
outdir  <- 'PRV-MDBIO-4cell'; try({ dir.create(outdir) })



#### Settings ####

## Main
#save.data <- 'PRV.rebase.RData'
EndType <- 'TSS'

#project_config  <- fread('project_config.txt')
#outdir  <- project_config$outdir
res.dir <- outdir; try({ dir.create(res.dir) })
fig.dir <- paste0(res.dir, '/', EndType, '_abund'); # try({ dir.create(fig.dir) })## save plots to this directory

## Miscallenaous
## Colors
colorvec <- na.omit(c(
  colorvec[c(1:5,15,13,10,24,7,26,32,33)],
  colorvec[-c(1:5,15,13,10,24,7,26,32,33)]))

colorvec <- colorvec[-c(9,13)]


colorvec <- colorvec[-c(10,13,17,18)]

colorvec <- colorvec[c(1:2,14,3:length(colorvec))]

palette <- colorvec  #pal_npg()(10)

##
writetables <- T
load.prev   <- F



## Filter metadata for those samples that had at least 1 viral read
readcounts <- fread(paste0(outdir, '/readcounts.tsv'), header = T)
metafilt   <- metafilt[metafilt$sample %in% readcounts$sample, ]


#### Normalization method
norm.method <- 'viral_read.count'
# Normalization before or after filtering for counts that didn't overlap with the other end?
norm_before_filtering <- F

```


# Project info
Samples: dcDNA Seq  
pychopper: no  
mapping version: v6  
LoRTIA: yes (stranded only output)  



# TO-DO:

## Integrate NRK cell line's other hpi samples
- metadata ✔  
- LoRTIA ✔  
- bamfiles ✔  
- GFF-compare -->> select not the best hit from the all results but the canonic !  ✔
- TransFrags ✔
- TSSr ✔
- prime plots
-->> everything, basically...


## Visualization
- Logarithmic scales for gene-pair ratios
- else ?

## Ref Transcripts
- Correct REF annotation's TES -> there were some very close ->> Gábor & Ádám  ✔
- Add canonic (T,F), gene, and CDS to REF transcripts  ->> Gábor & Ádám  ✔

## Transcript Generation  
- Venn diagrams -> how many reads have correct TES or TSS? -> this makes sense only sample-wise...
  --> check if there are TransFrags that have no correct TSS or TES, or rather
      --> make a distribution of in how many sample groups correct prime3 and prime5 ends were found for each




```{r}
#
CAGE_TSSr_clusters      <- fread(paste0(config$outdir,  '/TSSr.TSS.CAGE.all.Cluster.Results.txt'))
#
dcDNA_TSSr_TSS_clusters <- fread(paste0(config$outdir,  '/TSSr.TSS.dcDNA.all.Cluster.Results.txt'))
#
dcDNA_TSSr_TES_clusters <- fread(paste0(config$outdir,  '/TSSr.TES.dcDNA.all.Cluster.Results.txt'))
#
dRNA_TSSr_TES_clusters  <- fread(paste0(config$outdir,  '/TSSr.TES.dRNA.all.Cluster.Results.txt'))

outdir <- config$outdir


#### dcDNA TSS TSSr
TR.Ref.data.TSS.dcDNA        <- fread(paste0(outdir, '/TR.Ref.data.TSS.dcDNA.tsv'),        na.strings = '')
TR.gff.compare.uni.TSS.dcDNA <- fread(paste0(outdir, '/TR.gff.compare.uni.TSS.dcDNA.tsv'), na.strings = '')
TSS.dcDNA.support.freq       <- fread(paste0(outdir, '/TSS.dcDNA.support.freq.tsv'),       na.strings = '')
TSS.dcDNA_TSSr_clusters_uni  <- fread(paste0(outdir, '/TSS.dcDNA.TSSr_clusters_uni.tsv'),  na.strings = '')

TSS.dcDNA_ClusterShapes      <- fread(paste0(outdir, '/TSSr.TSS.dcDNA.all.ClusterShapes.txt'))
##
TSS.dcDNA_assignedClusters   <- fread(paste0(outdir, '/TSSr.TSS.dcDNA.best.assignedClusters.txt'), na.strings = '')


#### dcDNA TES TSSr
TR.Ref.data.TES.dcDNA        <- fread(paste0(outdir, '/TR.Ref.data.TES.dcDNA.tsv'),        na.strings = '')
TR.gff.compare.uni.TES.dcDNA <- fread(paste0(outdir, '/TR.gff.compare.uni.TES.dcDNA.tsv'), na.strings = '')
TES.dcDNA.support.freq       <- fread(paste0(outdir, '/TES.dcDNA.support.freq.tsv'),       na.strings = '')
TES.dcDNA_TSSr_clusters_uni  <- fread(paste0(outdir, '/TES.dcDNA.TSSr_clusters_uni.tsv'),  na.strings = '')

TES.dcDNA_ClusterShapes      <- fread(paste0(outdir, '/TSSr.TES.dcDNA.all.ClusterShapes.txt'))
##
TES.dcDNA_assignedClusters   <- fread(paste0(outdir, '/TSSr.TES.dcDNA.best.assignedClusters.txt'), na.strings = '')


#### dRNA TES TSSr
TR.Ref.data.TES.dRNA        <- fread(paste0(outdir, '/TR.Ref.data.TES.dRNA.tsv'),        na.strings = '')
TR.gff.compare.uni.TES.dRNA <- fread(paste0(outdir, '/TR.gff.compare.uni.TES.dRNA.tsv'), na.strings = '')
TES.dRNA.support.freq       <- fread(paste0(outdir, '/TES.dRNA.support.freq.tsv'),       na.strings = '')
TES.dRNA_TSSr_clusters_uni  <- fread(paste0(outdir, '/TES.dRNA.TSSr_clusters_uni.tsv'),  na.strings = '')

TES.dRNA_ClusterShapes      <- fread(paste0(outdir, '/TSSr.TES.dRNA.all.ClusterShapes.txt'))
##
TES.dRNA_assignedClusters   <- fread(paste0(outdir, '/TSSr.TES.dRNA.best.assignedClusters.txt'), na.strings = '')



#### CAGE TSSr
TR.Ref.data.CAGE        <- fread(paste0(outdir, '/TR.Ref.data.CAGE.tsv'),        na.strings = '')
TR.gff.compare.uni.CAGE <- fread(paste0(outdir, '/TR.gff.compare.uni.CAGE.tsv'), na.strings = '')
CAGE.support.freq       <- fread(paste0(outdir, '/CAGE.support.freq.tsv'),       na.strings = '')
CAGE_TSSr_clusters_uni  <- fread(paste0(outdir, '/CAGE.TSSr_clusters_uni.tsv'),  na.strings = '')

####
EX.uni    <- fread(paste0(outdir, '/EX.uni.tsv'),  na.strings = '')

#### dcDNA TransFrags Results
TR.gff.compare.merged.TR.counts.gt <- fread(paste0(outdir, "/TR.gff.compare.merged.TR.counts.gt.tsv.gz"), na.strings = '')

cols_to_group <- setdiff(colnames(TR.gff.compare.merged.TR.counts.gt), c(metacols, "count", 'rep', 'total', 'contig_size', "average_coverage", "norm_count"))
TR.gff.compare.uni <- unique(TR.gff.compare.merged.TR.counts.gt[,..cols_to_group])
dup(TR.gff.compare.uni$transcript_id)

TR.gff.compare.uni[,prime5 := fifelse(strand == '+', start, end)]
TR.gff.compare.uni[,prime3 := fifelse(strand == '+', end, start)]

TR.gff.compare.uni[,TR.prime5.win.start := prime5]
TR.gff.compare.uni[,TR.prime5.win.end   := prime5]
TR.gff.compare.uni[,TR.prime3.win.start := prime3]
TR.gff.compare.uni[,TR.prime3.win.end   := prime3]


```

# TSSr Clustering results of CAGE and dcDNA data

```{r, fig.width=32, fig.height=6, eval=F}
dcDNA_TSSr_clusters_uni[,TC.width := (TC.end - TC.start) + 1]


## histogram
ggplot(dcDNA_TSSr_clusters_uni) +
  geom_rect(#data = dcDNA_TSSr_clusters_uni[strand == '+'],
            aes(xmin = TC.start, xmax = TC.end, ymin = 1, ymax = 2)) +
  #xlim(c(0,250))  +
  facet_nested(rows=vars(strand)) +
  theme_minimal() +
  theme(legend.position = 'bottom')


```

## Cluster Widths

### dcDNA TSS
```{r}
TSS.dcDNA_TSSr_clusters_uni[,TC.width := (TC.end - TC.start) + 1]


## histogram
ggplot(TSS.dcDNA_TSSr_clusters_uni) +
  geom_histogram(aes(TC.width), binwidth =  10) +
  #xlim(c(0,250))  +
  #facet_nested(rows=vars(hpi)) +
  theme_minimal() +
  theme(legend.position = 'bottom')


## histogram
ggplot(TSS.dcDNA_TSSr_clusters_uni) +
  geom_bar(aes(TC.width)) +
  xlim(c(0,50))  +
  #facet_nested(rows=vars(hpi)) +
  theme_minimal() +
  theme(legend.position = 'bottom')

```

### dcDNA TES
```{r}
TES.dcDNA_TSSr_clusters_uni[,TC.width := (TC.end - TC.start) + 1]


## histogram
ggplot(TES.dcDNA_TSSr_clusters_uni) +
  geom_histogram(aes(TC.width), binwidth =  10) +
  #xlim(c(0,250))  +
  #facet_nested(rows=vars(hpi)) +
  theme_minimal() +
  theme(legend.position = 'bottom')


## histogram
ggplot(TES.dcDNA_TSSr_clusters_uni) +
  geom_bar(aes(TC.width)) +
  xlim(c(0,50))  +
  #facet_nested(rows=vars(hpi)) +
  theme_minimal() +
  theme(legend.position = 'bottom')

```

### CAGE
```{r}
CAGE_TSSr_clusters_uni[,TC.width := (TC.end - TC.start) + 1]


## histogram
ggplot(CAGE_TSSr_clusters_uni) +
  geom_histogram(aes(TC.width), binwidth =  10) +
  #xlim(c(0,250))  +
  #facet_nested(rows=vars(hpi)) +
  theme_minimal() +
  theme(legend.position = 'bottom')

## histogram
ggplot(CAGE_TSSr_clusters_uni) +
  geom_bar(aes(TC.width), binwidth =  10) +
  xlim(c(0,50))  +
  #facet_nested(rows=vars(hpi)) +
  theme_minimal() +
  theme(legend.position = 'bottom')

```

## Assigning clutsers to genes

```{r}
```


### Plot top20 Gene-Wise Shifts TSS
```{r, fig.width=12, fig.height=12}
gene_TSS_shifts_Sum <- unique(TSS.dcDNA_assignedClusters[,
                                .(chr, strand, gene, gene_mean_shift_TSS, gene_sd_shift_TSS, gene_varcoeff_shif_TSS)])
gene_TSS_shifts_Sum <- gene_TSS_shifts_Sum[order(gene_mean_shift_TSS, decreasing = T)]
topShifts   <- gene_TSS_shifts_Sum[1:20, gene]
topShifts   <- topShifts[!topShifts %in% c('US1_2', 'IE180_2')]

### Plot Gene-Wise Shifts
ggplot(TSS.dcDNA_assignedClusters[gene %in% topShifts
                      & Time %in% c(1,2,4,6,8,12)  ], 
       aes(Time, gene_dominant_TSS_shift)) + 
  geom_line(aes(color = cell_line, group = cell_line)) + 
  geom_point(aes(color = cell_line)) + 
  #coord_flip() +
  facet_nested_wrap(~gene, ncol=4) +
  theme_minimal()



```

### Plot top20 Gene-Wise Shifts TES
```{r, fig.width=12, fig.height=12}
gene_TES_shifts_Sum <- unique(TES.dcDNA_assignedClusters[,
                                .(chr, strand, gene, gene_mean_shift_TES, gene_sd_shift_TES, gene_varcoeff_shif_TES)])
gene_TES_shifts_Sum <- gene_TES_shifts_Sum[order(gene_mean_shift_TES, decreasing = T)]
topShifts   <- gene_TES_shifts_Sum[1:20, gene]
topShifts   <- topShifts[!topShifts %in% c('US1_2', 'IE180_2')]

### Plot Gene-Wise Shifts
ggplot(TES.dcDNA_assignedClusters[gene %in% topShifts
                      & Time %in% c(1,2,4,6,8,12)  ], 
       aes(Time, gene_dominant_TES_shift)) + 
  geom_line(aes(color = cell_line, group = cell_line)) + 
  geom_point(aes(color = cell_line)) + 
  #coord_flip() +
  facet_nested_wrap(~gene, ncol=4) +
  theme_minimal()



```

### Statistical Comparison
```{r}

```


# Merge dcDNA TSSr Clustering results and CAGE TSSr clustering results with TransFrags and Reference Transcripts


## The Transcripts were merged based on their TSS with both the dcDNA-based TSS lusters and the CAGE TSS clusters


## Reference Transcripts
```{r}
#TR.Ref.data.TSSr.comb <- copy(TR.Ref.data.CAGE)
cnames <- c('TC.start', 'TC.end', 'support', 'score', 'consensus_peak')

## TSS
setnames(TR.Ref.data.CAGE,      cnames, paste0('CAGE.TSS_',  cnames))
setnames(TR.Ref.data.TSS.dcDNA, cnames, paste0('dcDNA.TSS_', cnames))

TR.Ref.data.TSSr.TSS.comb <- merge(TR.Ref.data.CAGE,
                                   TR.Ref.data.TSS.dcDNA,
                                   by=colnames(TR.Ref.data.CAGE)[2:17])

## TES
setnames(TR.Ref.data.TES.dRNA,  cnames, paste0('dRNA.TES_',  cnames))
setnames(TR.Ref.data.TES.dcDNA, cnames, paste0('dcDNA.TES_', cnames))

TR.Ref.data.TSSr.TES.comb <- merge(TR.Ref.data.TES.dcDNA,
                                   TR.Ref.data.TES.dRNA,
                                   by=colnames(TR.Ref.data.TES.dcDNA)[2:17])


ggVennDiagram::ggVennDiagram(list(
  #REF_TR = TR.Ref.data$transcript_id,
  
  CAGE_TSS_supported   = TR.Ref.data.CAGE[      !is.na(CAGE_TC_ID),  transcript_id],
  dcDNA_TSS_supported  = TR.Ref.data.TSS.dcDNA[ !is.na(dcDNA.TSS_TC_ID), transcript_id],
  dcDNA.TES_supported  = TR.Ref.data.TES.dcDNA[ !is.na(dcDNA.TES_TC_ID), transcript_id],
  dRNA.TES_supported   = TR.Ref.data.TES.dRNA [ !is.na(dRNA.TES_TC_ID),  transcript_id]
  
  ))
```


## TransFrags
```{r}

### TransFrags
#TR.Ref.data.TSSr.comb <- copy(TR.Ref.data.CAGE)
cnames <- c('TC.start', 'TC.end', 'support', 'score', 'consensus_peak')

## TSS
setnames(TR.gff.compare.uni.CAGE,  cnames, paste0('CAGE.TSS_',  cnames))
setnames(TR.gff.compare.uni.TSS.dcDNA, cnames, paste0('dcDNA.TSS_', cnames))

TransFrags.TSSr.TSS.comb <- merge(TR.gff.compare.uni.CAGE, 
                                  TR.gff.compare.uni.TSS.dcDNA,
                                  by=colnames(TR.gff.compare.uni.CAGE)[c(2:4,10:27)])



## TES
setnames(TR.gff.compare.uni.TES.dRNA,  cnames, paste0('dRNA.TES_', cnames))
setnames(TR.gff.compare.uni.TES.dcDNA, cnames, paste0('dcDNA.TES_',  cnames))

TransFrags.TSSr.TES.comb <- merge( TR.gff.compare.uni.TES.dcDNA,
                                   TR.gff.compare.uni.TES.dRNA,
                                   by=colnames(TR.gff.compare.uni.TES.dcDNA)[c(2:4,10:27)])


ggVennDiagram::ggVennDiagram(list(
  #TransFrag           = TR.gff.compare.uni$transcript_id,

  CAGE_TSS_supported   = TR.gff.compare.uni.CAGE[      !is.na(CAGE_TC_ID),      transcript_id],
  dcDNA_TSS_supported  = TR.gff.compare.uni.TSS.dcDNA[ !is.na(dcDNA.TSS_TC_ID), transcript_id],
  dcDNA.TES_supported  = TR.gff.compare.uni.TES.dcDNA[ !is.na(dcDNA.TES_TC_ID), transcript_id],
  dRNA.TES_supported   = TR.gff.compare.uni.TES.dRNA [ !is.na(dRNA.TES_TC_ID),  transcript_id]

))
```


```{r}

## Merge TSS and TES

stopifnot(all.equal(colnames(TransFrags.TSSr.TSS.comb)[1:21], colnames(TransFrags.TSSr.TES.comb)[1:21])) 


colnames(TransFrags.TSSr.TES.comb)[-c(1:21)]
colnames(TransFrags.TSSr.TSS.comb)[-c(1:21)]



TransFrags.TSSr.comb <- merge(TransFrags.TSSr.TES.comb, 
                              TransFrags.TSSr.TSS.comb,
                              by = colnames(TransFrags.TSSr.TES.comb)[1:21] )


```

```{r}
### Non-spliced Transfrags ONly !!!
TR_EX           <- dcast(EX.uni, seqnames + strand + TR_ID ~ exon_rank, value.var = 'exon_ID')
non.spliced.TFs <- TR_EX[is.na(`2`), TR_ID]

TransFrags.TSSr.comb <- TransFrags.TSSr.comb[transcript_id %in% non.spliced.TFs, ]
```


## Validate TESs of TransFrags
```{r, eval = T}
### Validate TESs
#TransFrags.TSSr.comb[, TES   := prime3] # TransFrags.TSSr.comb[, TES   := NULL]

# "Valid" prime3 sites form reference transcripts
#valid.prime3[,TES := start + 10]
#fwrite(valid.prime3, 'valid.prime3.tsv', sep = '\t')

# had to correct, as some TESs were very close ! -->> cehck !!!
#valid.prime3 <- fread('valid.prime3.tsv')
# -> DONE

valid.TES.win <- 10

TR.ref[,transcript_prime3 := fifelse(strand == '+', transcript_end, transcript_start)]
TR.ref[,transcript_prime5 := fifelse(strand == '-', transcript_end, transcript_start)]

valid.prime3 <- unique(TR.ref[,.(seqnames, transcript_start, transcript_end, transcript_prime3, transcript_prime5, strand)])
valid.prime3 <- valid.prime3[,.(seqnames, strand, TES = transcript_prime3, start = transcript_prime3 - valid.TES.win, end = transcript_prime3 + valid.TES.win)]
valid.prime3 <- unique(valid.prime3)
valid.prime3[,valid_tes := T]

# make start and end as prime3 for joining
#TransFrags.TSSr.comb[,start := prime3]
#TransFrags.TSSr.comb[,end   := prime3]
prime3.TF.ov <- copy(TransFrags.TSSr.comb)
prime3.TF.ov[, start := prime3][, end := prime3]


prime3.TF.ov <- foverlaps2(prime3.TF.ov, valid.prime3,
                           by.x=c('seqnames', 'strand', 'start', 'end'),
                           by.y=c('seqnames', 'strand', 'start', 'end'), minoverlap = 1)


prime3.TF.ov <- unique(prime3.TF.ov[,.(seqnames, strand, transcript_id, valid_tes, TES)])

# Merge with TransFrags
TransFrags.TSSr.comb <- merge(TransFrags.TSSr.comb, prime3.TF.ov, by=c('seqnames', 'strand', 'transcript_id'), all.x=T)
TransFrags.TSSr.comb[is.na(valid_tes), valid_tes := F]

# correct start and end
#TransFrags.TSSr.comb[, start  := fifelse(strand == '+', prime5, prime3)]
#TransFrags.TSSr.comb[, end    := fifelse(strand == '+', prime3, prime5)]

prime3.valid.corr.freq <- TransFrags.TSSr.comb[,.N, by=.(valid_tes)]

ggvf <- ggplot(prime3.valid.corr.freq) +
  geom_col(aes(x=valid_tes, y=N, fill=valid_tes), color='black') +
  theme_bw() +
  ggtitle('TransFrags 3-prime end result, according to ref mRNA TES')

ggsave(file.path(outdir, 'TF_TES_validation.jpg'), ggvf, height = 12, width = 9)


ggvf
```


## Annotate new Transcripts from TSS-validated TransFrags {.tabset}


### TransFrags that are not equal to Reference TX
```{r}
ggVennDiagram::ggVennDiagram(list(
  
  CAGE_TSS_supported   = TransFrags.TSSr.comb[ !is.na(CAGE_TC_ID)      & class_code != '=' ,  transcript_id],
  dcDNA_TSS_supported  = TransFrags.TSSr.comb[ !is.na(dcDNA.TSS_TC_ID) & class_code != '=' ,  transcript_id],
  dcDNA.TES_supported  = TransFrags.TSSr.comb[ !is.na(dcDNA.TES_TC_ID) & class_code != '=' ,  transcript_id],
  dRNA.TES_supported   = TransFrags.TSSr.comb[ !is.na(dRNA.TES_TC_ID)  & class_code != '=' ,  transcript_id]
  
))
```

### TransFrags that are have at least '**' 
```{r}
ggVennDiagram::ggVennDiagram(list(
  
  CAGE_TSS_supported   = TransFrags.TSSr.comb[ !is.na(CAGE_TC_ID)      & class_code != '=' & CAGE_TC_significance       != '*' ,  transcript_id],
  dcDNA_TSS_supported  = TransFrags.TSSr.comb[ !is.na(dcDNA.TSS_TC_ID) & class_code != '=' & dcDNA.TSS_TC_significance  != '*' ,  transcript_id],
  dcDNA.TES_supported  = TransFrags.TSSr.comb[ !is.na(dcDNA.TES_TC_ID) & class_code != '=' & dcDNA.TES_TC_significance  != '*' ,  transcript_id],
  dRNA.TES_supported   = TransFrags.TSSr.comb[ !is.na(dRNA.TES_TC_ID)  & class_code != '=' & dRNA.TES_TC_significance   != '*' ,  transcript_id]
  
))
```

### TransFrags that share 3-prime with previous annotation -->> significant drop
```{r}
ggVennDiagram::ggVennDiagram(list(
  
  CAGE_TSS_supported   = TransFrags.TSSr.comb[ !is.na(CAGE_TC_ID)      & class_code != '=' & CAGE_TC_significance       != '*' & valid_tes == T,  transcript_id],
  dcDNA_TSS_supported  = TransFrags.TSSr.comb[ !is.na(dcDNA.TSS_TC_ID) & class_code != '=' & dcDNA.TSS_TC_significance  != '*' & valid_tes == T,  transcript_id],
  dcDNA.TES_supported  = TransFrags.TSSr.comb[ !is.na(dcDNA.TES_TC_ID) & class_code != '=' & dcDNA.TES_TC_significance  != '*' & valid_tes == T,  transcript_id],
  dRNA.TES_supported   = TransFrags.TSSr.comb[ !is.na(dRNA.TES_TC_ID)  & class_code != '=' & dRNA.TES_TC_significance   != '*' & valid_tes == T,  transcript_id]

))
```

### Filtering of TransFrags

*No filtering the TransFrags now, rather we will filter the final Transcripts*

```{r, eval=T}

filter.TFRs <- F

if(filter.TFRs) {
  TransFrags.TSSr.filt <-  TransFrags.TSSr.comb[  (!is.na(CAGE_TC_ID)  & CAGE_TC_significance   %in% c('**', '***'))
                                                | (!is.na(dcDNA_TC_ID) & dcDNA_TC_significance  %in% c('**', '***'))
                                                #& class_code != '='  & valid_tes == T
                                                ,  ]
  
  TransFrags.TSSr.filt[, .N, by = .(CAGE_TC_significance, dcDNA_TC_significance)]
  
  TransFrags.TSSr.filt <-  TransFrags.TSSr.filt[class_code != '='  & valid_tes == T, ]
  
  TransFrags.TSSr.filt[, .N, by = .(CAGE_TC_significance, dcDNA_TC_significance, valid_tes)]
  
} else {
  
  TransFrags.TSSr.filt <- TransFrags.TSSr.comb
  
}


```

```{r, eval=F}
TransFrags.TSSr.filt.counts.gt[,.N,.(group,hpi,cell_line)]
dRNA_TSSr_TES_clusters[,.N,.(group,cell_line)]
```

## Merge TransFrags with TSSr group-wise cluster results (for the peaks)
```{r}
### Merge TransFrags with TSSr group-wise cluster results (for the peaks)
TransFrags.TSSr.filt.counts.gt <- merge(TR.gff.compare.merged.TR.counts.gt[,.(seqnames, strand, transcript_id, sample, rep, group, cell_line, hpi, Time, count, norm_count)], 
                                        TransFrags.TSSr.filt,
                                        by = c('seqnames', 'strand', 'transcript_id'),
                                        all.y = T)

TransFrags.TSSr.filt.counts.gt[, cell_line := gsub('-', '_', cell_line)]
TransFrags.TSSr.filt.counts.gt[, group     := gsub('-', '_', group)]

### Columns to rename
cnames <- c('cluster.start', 'cluster.end', 'dominant_tss', 'tags', 'tags.dominant_tss',  'q_0.1', 'q_0.9', 'interquantile_width', 'shape.score', 'cluster_width')



### Merge with dcDNA TSS clusters
TransFrags.TSSr.filt.counts.gt  <- merge(TransFrags.TSSr.filt.counts.gt, 
                                         dcDNA_TSSr_TSS_clusters[,.(seqnames = chr, strand, cell_line, hpi, 
                                                                  cluster.start = start, cluster.end = end, dcDNA.TSS_TC_ID = paste0('TC_', cluster),
                                                                  dominant_tss, tags, tags.dominant_tss,  q_0.1, q_0.9, interquantile_width, shape.score, cluster_width)],
                                         #by.y = c('seqnames', 'strand', 'cluster', 'group'),
                                         by = c('seqnames', 'strand', 'dcDNA.TSS_TC_ID', 'cell_line', 'hpi'),
                                         all.x = T )

# Calculate distance of TransFrags to peak of TC                      
TransFrags.TSSr.filt.counts.gt[, TSS_dcDNA_distance := dominant_tss - prime5, by= transcript_id]

# Rename columns
setnames(TransFrags.TSSr.filt.counts.gt,  cnames, paste0('dcDNA.TSS_', cnames))


### Merge with CAGE clusters
TransFrags.TSSr.filt.counts.gt  <- merge(TransFrags.TSSr.filt.counts.gt, 
                                         CAGE_TSSr_clusters[,.(seqnames = chr, strand, cell_line, hpi, 
                                                                cluster.start = start, cluster.end = end, CAGE_TC_ID = paste0('TC_', cluster),
                                                                dominant_tss, tags, tags.dominant_tss,  q_0.1, q_0.9, interquantile_width, shape.score, cluster_width)],
                                         #by.y = c('seqnames', 'strand', 'cluster', 'group'),
                                         by = c('seqnames', 'strand', 'CAGE_TC_ID', 'cell_line', 'hpi'),
                                         all.x = T )

# Calculate distance of TransFrags to peak of TC                      
TransFrags.TSSr.filt.counts.gt[, TSS_CAGE_distance := dominant_tss - prime5, by= transcript_id]

# Rename columns
setnames(TransFrags.TSSr.filt.counts.gt,  cnames, paste0('CAGE_', cnames))



### Merge with dcDNA TES clusters
TransFrags.TSSr.filt.counts.gt  <- merge(TransFrags.TSSr.filt.counts.gt, 
                                         dcDNA_TSSr_TES_clusters[,.(seqnames = chr, strand, cell_line, hpi, 
                                                                  cluster.start = start, cluster.end = end, dcDNA.TES_TC_ID = paste0('TC_', cluster),
                                                                  dominant_tes, tags, tags.dominant_tes,  q_0.1, q_0.9, interquantile_width, shape.score, cluster_width)],
                                         #by.y = c('seqnames', 'strand', 'cluster', 'group'),
                                         by = c('seqnames', 'strand', 'dcDNA.TES_TC_ID', 'cell_line', 'hpi'),
                                         all.x = T )

# Caluclate distance of TransFrags to peak of TC                      
TransFrags.TSSr.filt.counts.gt[, TES_dcDNA_distance := dominant_tes - prime3, by= transcript_id]

# Rename columns
cnames <- c('cluster.start', 'cluster.end', 'dominant_tes', 'tags', 'tags.dominant_tes',  'q_0.1', 'q_0.9', 'interquantile_width', 'shape.score', 'cluster_width')
setnames(TransFrags.TSSr.filt.counts.gt,  cnames, paste0('dcDNA.TES_', cnames))


## Merge with dRNA TES clusters
TransFrags.TSSr.filt.counts.gt  <- merge(TransFrags.TSSr.filt.counts.gt, 
                                         dRNA_TSSr_TES_clusters[,.(seqnames = chr, strand, group,
                                                                  cluster.start = start, cluster.end = end, dRNA.TES_TC_ID = paste0('TC_', cluster),
                                                                  dominant_tes, tags, tags.dominant_tes,  q_0.1, q_0.9, interquantile_width, shape.score, cluster_width)],
                                         #by.y = c('seqnames', 'strand', 'cluster', 'group'),
                                         by = c('seqnames', 'strand', 'dRNA.TES_TC_ID', 'group'),
                                         all.x = T )

# Caluclate distance of TransFrags to peak of TC                      
TransFrags.TSSr.filt.counts.gt[, TES_dRNA_distance := dominant_tes - prime3, by= transcript_id]

# Rename columns
cnames <- c('cluster.start', 'cluster.end', 'dominant_tes', 'tags', 'tags.dominant_tes',  'q_0.1', 'q_0.9', 'interquantile_width', 'shape.score', 'cluster_width')
setnames(TransFrags.TSSr.filt.counts.gt,  cnames, paste0('dRNA.TES_', cnames))

```


```{r}
# These distances are from the group-wise clusters

# Calculate the distance from consensus peak also
TransFrags.TSSr.filt.counts.gt[,   
     CAGE.TSS_consensus_peak_dist := CAGE.TSS_consensus_peak  - prime5
][, dcDNA.TSS_consensus_peak_dist := dcDNA.TSS_consensus_peak - prime5
][, dcDNA.TES_consensus_peak_dist := dcDNA.TES_consensus_peak - prime3
][,  dRNA.TES_consensus_peak_dist := dRNA.TES_consensus_peak  - prime3
  ]


```


### Distribution of Transfrags in the sample groups
```{r, fig.width= 12, fig.height=16}
ggplot(TransFrags.TSSr.filt.counts.gt) + 
  geom_col(aes(group, count, fill = cell_line)) +
  facet_grid(rows=vars(as.factor(Time)), scales = 'free') +
  coord_flip()
  
```



### Distance to dcDNA TSS cluster's group dominant TSS (peak)
```{r}
# Check distances
TransFrags.TSSr.filt.counts.gt[, .N, TSS_dcDNA_distance]
ggplot(TransFrags.TSSr.filt.counts.gt, aes(TSS_dcDNA_distance) ) + geom_histogram(bins=100) + 
  geom_vline(xintercept = c(-10, 10), color='red') +
  geom_vline(xintercept = c(-25, 25))
```

### Distance to dcDNA TSS cluster's consensus TSS peak
```{r}
TransFrags.TSSr.filt.counts.gt[, .N, dcDNA.TSS_consensus_peak_dist]
ggplot(TransFrags.TSSr.filt.counts.gt, aes(dcDNA.TSS_consensus_peak_dist) ) + geom_histogram(bins=100) + 
  geom_vline(xintercept = c(-10, 10), color='red') +
  geom_vline(xintercept = c(-25, 25))

```



### Distance to CAGE cluster's dominant TSS (peak)
```{r}
TransFrags.TSSr.filt.counts.gt[, .N, TSS_CAGE_distance]
ggplot(TransFrags.TSSr.filt.counts.gt, aes(TSS_CAGE_distance) ) + geom_histogram(bins=100) + 
  geom_vline(xintercept = c(-10, 10), color='red') +
  geom_vline(xintercept = c(-25, 25))
```

### Distance to CAGE TSS cluster's consensus TSS peak
```{r}
TransFrags.TSSr.filt.counts.gt[, .N, CAGE.TSS_consensus_peak_dist]
ggplot(TransFrags.TSSr.filt.counts.gt, aes(CAGE.TSS_consensus_peak_dist) ) + geom_histogram(bins=100) + 
  geom_vline(xintercept = c(-10, 10), color='red') +
  geom_vline(xintercept = c(-25, 25))

```


### Distance to dcDNA TES cluster's group dominant TES (peak)
```{r}
# Check distances
TransFrags.TSSr.filt.counts.gt[, .N, TES_dcDNA_distance]
ggplot(TransFrags.TSSr.filt.counts.gt, aes(TES_dcDNA_distance) ) + geom_histogram(bins=100) + 
  geom_vline(xintercept = c(-10, 10), color='red') +
  geom_vline(xintercept = c(-25, 25))
```

### Distance to dcDNA TES cluster's consensus TES peak
```{r}
TransFrags.TSSr.filt.counts.gt[, .N, dcDNA.TES_consensus_peak_dist]
ggplot(TransFrags.TSSr.filt.counts.gt, aes(dcDNA.TES_consensus_peak_dist) ) + geom_histogram(bins=100) + 
  geom_vline(xintercept = c(-10, 10), color='red') +
  geom_vline(xintercept = c(-25, 25))

```


### Distance to dRNA TES cluster's group dominant TES (peak)
```{r}
# Check distances
TransFrags.TSSr.filt.counts.gt[, .N, TES_dRNA_distance]
ggplot(TransFrags.TSSr.filt.counts.gt, aes(TES_dRNA_distance) ) + geom_histogram(bins=100) + 
  geom_vline(xintercept = c(-10, 10), color='red') +
  geom_vline(xintercept = c(-25, 25))
```

### Distance to dRNA TES cluster's consensus TES peak
```{r}
TransFrags.TSSr.filt.counts.gt[, .N, dRNA.TES_consensus_peak_dist]
ggplot(TransFrags.TSSr.filt.counts.gt, aes(dRNA.TES_consensus_peak_dist) ) + geom_histogram(bins=100) + 
  geom_vline(xintercept = c(-10, 10), color='red') +
  geom_vline(xintercept = c(-25, 25))

```





## TransFrags that share 5-primes with CAGE or dcDNA TSS cluster peak in the respective sample group with a +/- 10 bp wobble
```{r}
## TransFrags that share 5-primes with CAGE or dcDNA TSS cluster peak in the respective sample group with a +/- 10 bp wobble
ggVennDiagram::ggVennDiagram(list(
  
  CAGE_TSS_supported   = unique(TransFrags.TSSr.filt.counts.gt[abs(TSS_CAGE_distance)  <= 10 | abs(CAGE.TSS_consensus_peak_dist)  <= 10,  transcript_id]),
  dcDNA_TSS_supported  = unique(TransFrags.TSSr.filt.counts.gt[abs(TSS_dcDNA_distance) <= 10 | abs(dcDNA.TSS_consensus_peak_dist) <= 10,  transcript_id]),
  dcDNA_TES_supported  = unique(TransFrags.TSSr.filt.counts.gt[abs(TES_dcDNA_distance) <= 10 | abs(dcDNA.TES_consensus_peak_dist) <= 10,  transcript_id]),
  dRNA_TES_supported   = unique(TransFrags.TSSr.filt.counts.gt[abs(TES_dRNA_distance)  <= 10 | abs(dRNA.TES_consensus_peak_dist)  <= 10,  transcript_id])
  
))
```


```{r, eval=T}
peak.TSS.dist.thresh <- 10
peak.TES.dist.thresh <- 10


gg <- ggVennDiagram::Venn(list(
  
  CAGE_TSS_supported   = unique(TransFrags.TSSr.filt.counts.gt[!is.na(TSS_CAGE_distance)  & abs(TSS_CAGE_distance)   <= peak.TSS.dist.thresh,  transcript_id]),
  dcDNA_TSS_supported  = unique(TransFrags.TSSr.filt.counts.gt[!is.na(TSS_dcDNA_distance) & abs(TSS_dcDNA_distance)  <= peak.TSS.dist.thresh,  transcript_id]),
  dcDNA_TES_supported  = unique(TransFrags.TSSr.filt.counts.gt[!is.na(TES_dcDNA_distance) & abs(TES_dcDNA_distance)  <= peak.TES.dist.thresh,  transcript_id]),
  dRNA_TES_supported   = unique(TransFrags.TSSr.filt.counts.gt[!is.na(TES_dRNA_distance)  & abs(TES_dRNA_distance)   <= peak.TES.dist.thresh,  transcript_id])
  
))


transfrags.to.keep <- unique(c(
  ggVennDiagram::overlap(gg, c(1,3)), ## CAGE  TSS and dcDNA TES
  ggVennDiagram::overlap(gg, c(1,4)), ## CAGE  TSS and dRNA  TES
  ggVennDiagram::overlap(gg, c(2,3)), ## dcDNA TSS and dcDNA TES
  ggVennDiagram::overlap(gg, c(2,4))  ## dcDNA TSS and dRNA  TES
))

#TransFrags.TSSr.filt.counts.gt <- TransFrags.TSSr.filt.counts.gt[  ( abs(TSS_CAGE_distance)  <= 10 | abs(TSS_dcDNA_distance) <= 10 )
#                                                                 & ( abs(TES_dcDNA_distance) <= 10 | abs(TES_dRNA_distance)  <= 10 )
#                                                                   , ]

TransFrags.TSSr.filt.counts.gt <- TransFrags.TSSr.filt.counts.gt[transcript_id %in% transfrags.to.keep, ]

```



```{r, eval=F}
peak.TSS.dist.thresh <- 10
peak.TES.dist.thresh <- 10


transfrags.to.keep <- unique(TransFrags.TSSr.filt.counts.gt[ 
  ( (!is.na(TSS_CAGE_distance)  & abs(TSS_CAGE_distance)  <= peak.TSS.dist.thresh) | 
    (!is.na(TSS_dcDNA_distance) & abs(TSS_dcDNA_distance) <= peak.TSS.dist.thresh) ) &
  ( (!is.na(TES_dcDNA_distance) & abs(TES_dcDNA_distance) <= peak.TES.dist.thresh) | 
    (!is.na(TES_dRNA_distance)  & abs(TES_dRNA_distance)  <= peak.TES.dist.thresh) )
, transcript_id])



TransFrags.TSSr.filt.counts.gt <- TransFrags.TSSr.filt.counts.gt[transcript_id %in% transfrags.to.keep, ]

```


```{r}
ggplot(TransFrags.TSSr.filt.counts.gt) + 
  geom_col(aes(group, count, fill = cell_line)) +
  facet_grid(rows=vars(as.factor(Time)), #cols=vars(passes_filter), 
             scales = 'free') +
  coord_flip()

```


```{r, eval = F}
TSS.dRNA  <- unique(TransFrags.TSSr.filt.counts.gt[grep('_dRNA', group), .(seqnames, strand, prime5)])

TSS.dcDNA <- unique(TransFrags.TSSr.filt.counts.gt[grep('_dRNA', group, invert = T), .(seqnames, strand, prime5)])

TSS.ov <- foverlaps2( 
  TSS.dcDNA[,.(seqnames, strand, start = prime5 - 10, end = prime5 + 10)],
  TSS.dRNA[ ,.(seqnames, strand, start = prime5, end = prime5)],
  minoverlap = 1)


TransFrags.TSSr.filt.counts.gt[group == 'PK_15_dRNA']
```


```{r, eval = F}

TransFrags.TSSr.filt.counts.gt[is.na(TSS_CAGE_distance) & is.na(TSS_dcDNA_distance) & grep('dRNA', group, invert = T),
                               .(strand, transcript_id, group,
                                 TSS_CAGE_distance, TSS_dcDNA_distance, TES_dcDNA_distance, TES_dRNA_distance)]


TransFrags.TSSr.filt.counts.gt[transcript_id == 183746, ]
```



*The TSS here will thus be the Group-Wise DOMINANT peak of the dcDNA-based TC -- or, if thats missing, that of the Group-Wise DOMINANT CAGE peak,*  
*considering the +/- 10 bp wobble*


```{r, eval = T}
# The TSS here will thus be the peak of the TC (either in the CAGE ir dcDNA - which is closer)

TransFrags.TSSr.filt.counts.gt[, TSS := fifelse(is.na(TSS_dcDNA_distance), CAGE_dominant_tss, 
                                          fifelse(is.na(TSS_CAGE_distance),  dcDNA.TSS_dominant_tss,
                                            fifelse(abs(TSS_dcDNA_distance) < abs(TSS_CAGE_distance), dcDNA.TSS_dominant_tss, CAGE_dominant_tss   
                                                    ))), by=transcript_id]

## thos yielded TX with TSS closer than 10
```


```{r, eval = T}
# The TES here will thus be the peak of the TC (either in the dRNA ir dcDNA - which is closer)

TransFrags.TSSr.filt.counts.gt[, TES := fifelse(is.na(TES_dcDNA_distance), dRNA.TES_dominant_tes, 
                                          fifelse(is.na(TES_dRNA_distance),  dcDNA.TES_dominant_tes,
                                            fifelse(abs(TES_dcDNA_distance) < abs(TES_dRNA_distance), dcDNA.TES_dominant_tes, dRNA.TES_dominant_tes
                                                    ))), by=transcript_id]

## thos yielded TX with TSS closer than 10
```


*There are many TransFrags that had no TSS or TES in the given sample group*

```{r}
TransFrags.TSSr.filt.counts.gt[is.na(TSS), .N, ]
TransFrags.TSSr.filt.counts.gt[is.na(TES), .N, ]
```


*The TSS here will thus be the CONSENSUS peak of the dcDNA-based TC -- or, if thats missing, that of the CAGE-based CONSENSUS peak,*  
*considering the +/- 10 bp wobble*


```{r, eval = T}
# The TSS here will thus be the peak of the TC (either in the CAGE ir dcDNA - which is closer)

TransFrags.TSSr.filt.counts.gt[, TSS := fifelse(is.na(TSS), 
                                          fifelse(is.na(dcDNA.TSS_consensus_peak_dist), CAGE.TSS_consensus_peak, 
                                            fifelse(is.na(CAGE.TSS_consensus_peak_dist),  dcDNA.TSS_consensus_peak,
                                              fifelse(abs(dcDNA.TSS_consensus_peak_dist) < abs(CAGE.TSS_consensus_peak_dist), dcDNA.TSS_consensus_peak, CAGE.TSS_consensus_peak
                                                      ))), TSS), by=transcript_id]

## thos yielded TX with TSS closer than 10
```

```{r, eval = F}
# OR: The TSS here will thus be the peak of the dcDNA-based TC -- or, if thats missing, that of the CAGE-based TC

TransFrags.TSSr.filt.counts.gt[, TSS := fifelse(is.na(TSS_dcDNA_distance), CAGE_dominant_tss, dcDNA_dominant_tss), by=transcript_id]


TransFrags.TSSr.filt.counts.gt[,.N, by=.(TSS, TSS_dcDNA_distance, TSS_CAGE_distance)]


```


```{r, eval = T}
# The TES here will thus be the peak of the TC (either in the dRNA ir dcDNA - which is closer)

TransFrags.TSSr.filt.counts.gt[, TES := fifelse(is.na(TES),
                                          fifelse(is.na(dcDNA.TES_consensus_peak_dist), dRNA.TES_consensus_peak, 
                                            fifelse(is.na(dRNA.TES_consensus_peak_dist),  dcDNA.TES_consensus_peak,
                                              fifelse(abs(dcDNA.TES_consensus_peak_dist) < abs(dRNA.TES_consensus_peak_dist), dcDNA.TES_consensus_peak, dRNA.TES_consensus_peak
                                                      ))), TES), by=transcript_id]

## thos yielded TX with TSS closer than 10
```


```{r}
TransFrags.TSSr.filt.counts.gt[is.na(TSS), .N, ]
TransFrags.TSSr.filt.counts.gt[is.na(TES), .N, ]


#TransFrags.TSSr.filt.counts.gt[is.na(dcDNA.TSS_consensus_peak) & is.na(CAGE.TSS_consensus_peak) & is.na(dcDNA.TSS_dominant_tss) & is.na(CAGE_dominant_tss ), ]
```


### Assemble Transcripts based on TSS and TES
```{r}
### Assemble Transcripts based on TSS and TES

## Add Novel TX ID
TransFrags.TSSr.filt.counts.gt[, NOV_TX_ID := paste0('NOV_TX_', .GRP), by = .(seqnames, strand, TSS, TES)]
```


```{r, eval = F}
## 
TransFrags.TSSr.filt.group.counts <- unique(TransFrags.TSSr.filt.counts.gt[,.(seqnames, strand, start, end, TSS, TES, NOV_TX_ID, 
                                                                #TSS_dcDNA_distance, TSS_CAGE_distance,
                                                                transcript_id, cmp_ref, class_code, 
                                                                group, hpi, Time, cell_line, group, sample, count)])

TransFrags.TSSr.filt.group.counts <- TransFrags.TSSr.filt.group.counts[,.(sum_count = sum(count)), by=.(seqnames, strand, TSS, TES, NOV_TX_ID, group, cell_line, hpi, Time)]

```


```{r}
### Still, we need to de-replicate the TSSs

TSS.uni <- unique(TransFrags.TSSr.filt.counts.gt[,.(seqnames, strand, TSS)])


### Merge with previous annotation

valid.TSS.win <- 10

TR.ref[,transcript_prime3 := fifelse(strand == '+', transcript_end, transcript_start)]
TR.ref[,transcript_prime5 := fifelse(strand == '-', transcript_end, transcript_start)]

valid.prime5 <- unique(TR.ref[,.(seqnames, transcript_start, transcript_end, transcript_prime3, transcript_prime5, strand)])
valid.prime5 <- valid.prime5[,.(seqnames, strand, valid.TSS = transcript_prime5, start = transcript_prime5 - valid.TSS.win, end = transcript_prime5 + valid.TSS.win)]
valid.prime5 <- unique(valid.prime5)
valid.prime5[,valid_tss := T]



# make start and end as prime5 for joining
prime5.TF.ov <- copy(TSS.uni)
prime5.TF.ov[, start := TSS][, end := TSS]

prime5.TF.ov <- foverlaps2(prime5.TF.ov, valid.prime5,
                           by.x=c('seqnames', 'strand', 'start', 'end'),
                           by.y=c('seqnames', 'strand', 'start', 'end'), minoverlap = 1)


prime5.TF.ov <- unique(prime5.TF.ov[,.(seqnames, strand, TSS, valid.TSS, valid_tss)])


prime5.TF.ov <- prime5.TF.ov[!duplicated(TSS), ]
#dup(prime5.TF.ov$TSS)

# Merge
TransFrags.TSSr.filt.counts.gt <- merge(TransFrags.TSSr.filt.counts.gt, prime5.TF.ov, by=c('seqnames', 'strand', 'TSS'), all.x=T)
TransFrags.TSSr.filt.counts.gt[is.na(valid_tss), valid_tss := F]


TransFrags.TSSr.filt.counts.gt[, .N, valid_tss]



TransFrags.TSSr.filt.counts.gt[,TSS := fifelse(valid_tss, valid.TSS, TSS) ]
```

```{r}
## Add Novel TX ID, again
TransFrags.TSSr.filt.counts.gt[, NOV_TX_ID := paste0('NOV_TX_', .GRP), by = .(seqnames, strand, TSS, TES)]

```




```{r}
### There are still some duplicate TSSs (closer than 10 bp) -->>

TSS.uni <- unique(TransFrags.TSSr.filt.counts.gt[,.(seqnames, strand, TSS)])
TSS.uni[, start := TSS][, end := TSS]

# Create group identifier: within each (seqnames, strand), if the difference between consecutive TSSs is less than 10,
# they are in the same group; otherwise start a new group.
TSS.uni[, grp := cumsum(c(TRUE, diff(TSS) >= valid.TSS.win)), by = .(seqnames, strand)]
# Now create a new column with the merged TSS value for each group
TSS.uni[, mergedTSS := round(mean(TSS), 0), by = .(seqnames, strand, grp)]
length(unique(TSS.uni[, mergedTSS]))

TransFrags.TSSr.filt.counts.gt <- merge(TransFrags.TSSr.filt.counts.gt, TSS.uni[,.(seqnames, strand, TSS, mergedTSS)], by=c('seqnames', 'strand', 'TSS') )

TransFrags.TSSr.filt.counts.gt[, TSS := mergedTSS ]
TransFrags.TSSr.filt.counts.gt[, mergedTSS := NULL]

```

```{r}
## Add Novel TX ID, again
TransFrags.TSSr.filt.counts.gt[, NOV_TX_ID := paste0('NOV_TX_', .GRP), by = .(seqnames, strand, TSS, TES)]

```



```{r}

### Finally, summarise the Nov TXs

TransFrags.TSSr.filt.group.counts <- TransFrags.TSSr.filt.counts.gt[,.(count = sum(count)), by=.(seqnames, strand, TSS, TES, NOV_TX_ID, group, cell_line, hpi, Time)]

TransFrags.TSSr.filt.group.counts[, prevalence := uniqueN(group), by = NOV_TX_ID]
TransFrags.TSSr.filt.group.counts[, sum_count  := sum(count),     by = NOV_TX_ID]
```


```{r}

### Prevalence Distribution

ggplot(TransFrags.TSSr.filt.group.counts) +
  geom_bar(aes(prevalence))

```

*Filtering for read count : >= 5*  
*Filtering for group prevalence : >= 2*  


```{r}
## Filter for read_count ?
novTX.readcount.filt  <- 5


## Filter for prevalence?
novTX.prevalence.filt <- 2

##
message('Keeping ', length(unique(TransFrags.TSSr.filt.group.counts[ prevalence >= novTX.prevalence.filt & sum_count >= novTX.readcount.filt, NOV_TX_ID])), 
        ' out of ', length(unique(TransFrags.TSSr.filt.group.counts[ , NOV_TX_ID])),
        ' Novel Transcripts (', 
        round((length(unique(TransFrags.TSSr.filt.group.counts[ prevalence >= novTX.prevalence.filt & sum_count >= novTX.readcount.filt, NOV_TX_ID])) / 
                 length(unique(TransFrags.TSSr.filt.group.counts[ , NOV_TX_ID]))) * 100, 0),
        '%)' )
```


```{r}
TransFrags.TSSr.filt.group.counts <- TransFrags.TSSr.filt.group.counts[ prevalence  >= novTX.prevalence.filt
                                                                      & sum_count   >= novTX.readcount.filt, ]

```


```{r}
TransFrags.TSSr.filt.group.counts[,start := fifelse(strand == '+', TSS, TES)]
TransFrags.TSSr.filt.group.counts[,end   := fifelse(strand == '+', TES, TSS)]


TransFrags.TSSr.filt.counts.gt[,start := fifelse(strand == '+', TSS, TES)]
TransFrags.TSSr.filt.counts.gt[,end   := fifelse(strand == '+', TES, TSS)]

## Filter melted table
TransFrags.TSSr.filt.counts.gt <- TransFrags.TSSr.filt.counts.gt[NOV_TX_ID %in% TransFrags.TSSr.filt.group.counts[,NOV_TX_ID], ]

## Unique Novel TXs (filtered)
NOV.TXs.uni    <- TransFrags.TSSr.filt.group.counts[,.(sum_count = sum(sum_count)), by = .(seqnames, strand, TSS, TES, NOV_TX_ID) ]

NOV.TXs.counts <- TransFrags.TSSr.filt.group.counts[,.(sum_count = sum(sum_count)), by = .(seqnames, strand, TSS, TES, NOV_TX_ID, group, hpi, Time, cell_line) ] 

NOV.TXs.uni.sp <- dcast(TransFrags.TSSr.filt.group.counts, seqnames + strand + TSS + TES + NOV_TX_ID ~ group, value.var = 'count', fill=0)

```


```{r}

ggplot(TransFrags.TSSr.filt.counts.gt) + 
  geom_col(aes(group, count, fill = cell_line)) +
  facet_grid(rows=vars(as.factor(Time)), 
             #cols=vars(passes_filter), 
             scales = 'free') +
  coord_flip()


```


```{r}
## TransFrags that share 5-primes with CAGE or dcDNA TSS cluster peak in the respective sample group with a +/- 10 bp wobble
ggVennDiagram::ggVennDiagram(list(
  PK_15           = unique(TransFrags.TSSr.filt.group.counts[ cell_line == 'PK_15', NOV_TX_ID]),
  PC_12           = unique(TransFrags.TSSr.filt.group.counts[ cell_line == 'PC_12', NOV_TX_ID]),
  C6              = unique(TransFrags.TSSr.filt.group.counts[ cell_line == 'C6',    NOV_TX_ID]),
  NRK             = unique(TransFrags.TSSr.filt.group.counts[ cell_line == 'NRK',   NOV_TX_ID])
))

```


### Include Adapter Counts to Novel Transcripts

```{r, eval=FALSE}


adapt.TR.sp <- fread(file.path(outdir, "adapt.TR.sp.tsv.gz"), sep = '\t')

## Import results back
TR.gff.compare.merged.TR.counts.gt <- fread(paste0(outdir, "/TR.gff.compare.merged.TR.counts.gt.tsv.gz"), na.strings = '')

### merge TransFrag count with Reference Transcript annotation for each Transfrag
TR.gff.compare.merged.TR.counts.gt <- merge(TR.gff.compare.merged.TR.counts.gt, adapt.TR.sp, by.x=c('transcript_id',  'sample'), by.y=c('TR_ID', 'sample') ) #, all = T)



###
TR.gff.compare.merged.TR.counts.gt.NovTX <- merge(TR.gff.compare.merged.TR.counts.gt, 
                                            unique(TransFrags.NovTX.uni[,.(seqnames, strand, transcript_id, NOV_TX_ID, TSS, TES)]),
                                            by = c('seqnames', 'strand', 'transcript_id'))




## sum counts again, this time really the counts. All counts combined (regardless of adapters). 
NOV.TXs.uni    <- TR.gff.compare.merged.TR.counts.gt.NovTX[,.(sum_count = sum(count)), by = .(seqnames, strand, TSS, TES, NOV_TX_ID) ]


NOV.TXs.counts <- TR.gff.compare.merged.TR.counts.gt.NovTX[,.(sum_count = sum(count)), by = .(seqnames, strand, TSS, TES, NOV_TX_ID, group, hpi, Time, cell_line) ] 

NOV.TXs.uni.sp <- dcast(NOV.TXs.counts, seqnames + strand + TSS + TES + NOV_TX_ID ~ group, value.var = 'sum_count', fill=0)

```


```{r}

ggplot(NOV.TXs.counts, aes(as.factor(Time), sum_count)) + 
  geom_col(aes(fill = cell_line), position = 'dodge')


ggplot(NOV.TXs.counts, aes(as.factor(Time), sum_count)) + 
  geom_col(aes(fill = cell_line), position = 'fill')


```


```{r}
fwrite(NOV.TXs.uni.sp, paste0(outdir, "/NOV.TXs.uni.sp.tsv"), sep = '\t')

fwrite(TransFrags.TSSr.filt.counts.gt, paste0(outdir, "/TransFrags.TSSr.filt.counts.gt.tsv.gz"), sep = '\t')
```


```{r}
NOV.TXs.uni    <- TransFrags.TSSr.filt.group.counts[,.(sum_count = sum(sum_count)), by = .(seqnames, strand, TSS, TES, NOV_TX_ID) ]

## GFF-formatted output
NOV.TXs.gff <- copy(NOV.TXs.uni)

NOV.TXs.gff[,start  := fifelse(strand == '+', TSS, TES)]
NOV.TXs.gff[,end    := fifelse(strand == '+', TES, TSS)]
NOV.TXs.gff[,type   := 'transcript']
NOV.TXs.gff[,source := 'Kakuk_et_al_2025']
NOV.TXs.gff[,score  := sum_count]
NOV.TXs.gff[,phase  := 0]
NOV.TXs.gff[,ID     := NOV_TX_ID]
NOV.TXs.gff[,Name   := NOV_TX_ID]


rtracklayer::export.gff3(NOV.TXs.gff, paste0(outdir, '/NOV.TXs.gff3'))
rtracklayer::export.gff3(NOV.TXs.gff, paste0(outdir, '/NOV.TXs.gff2'))
```

### Include Gene from Reference Transcripts


### Include TES from TSSr's TES analysis


### Use Gff-compare to compare Novel TXs to ALL REFERENCE Transcripts and select the CLOSEST REFERENCE
```{r}
###
ref_mRNAs <- unique(viral.mrna[,transcript_id])
#
out_dir <-paste0(outdir, '/gffcompare_iterative_NovTX')
#
all.gff.compare.outfile <- "NovTXs.merged.result_gff.compare.tsv.gz"
#
query_gff <- paste0(outdir, '/NOV.TXs.gff2')
#
config$compare.transfrags.to <- 'closest.Ref'


source('NovTX.gffcompare.R')
```


```{r}
## Import results back
best.merged.result_gff.compare <- fread(paste0(outdir, "/best.merged.result_gff.compare.NovTX.tsv.gz"), na.strings = '')


best.merged.result_gff.compare[,.N,class_code]


NOV.TXs.gff <- merge(NOV.TXs.gff, unique(best.merged.result_gff.compare[,.(seqnames, strand, cmp_ref, transcript_id, class_code)]),
                     by.x = c('seqnames', 'strand', 'NOV_TX_ID'), by.y = c('seqnames', 'strand', 'transcript_id'), all.x = T )

NOV.NOV.TXs.gff <- NOV.TXs.gff[class_code != '=' , ]


rtracklayer::export.gff3(NOV.NOV.TXs.gff, paste0(outdir, '/NOV.NOV.TXs.v2.gff3'))
rtracklayer::export.gff3(NOV.NOV.TXs.gff, paste0(outdir, '/NOV.NOV.TXs.v2.gff2'))


NOV.NOV.TXs.uni.sp <- NOV.TXs.uni.sp[NOV_TX_ID %in% NOV.NOV.TXs.gff$NOV_TX_ID, ]

fwrite(NOV.NOV.TXs.uni.sp, paste0(outdir, "/NOV.NOV.TXs.uni.sp.tsv"), sep = '\t')
```

```{r}

### Run comparing finction again, this time only on the canonical transcripts

###
ref_mRNAs <- unique(viral.mrna[,transcript_id])
#
out_dir <-paste0(outdir, '/gffcompare_iterative_Nov.NovTX')
#
all.gff.compare.outfile <- "Nov.NovTXs.merged.result_gff.compare.tsv.gz"
#
query_gff <- paste0(outdir, '/NOV.NOV.TXs.v2.gff2')
#
config$compare.transfrags.to <- 'canonic.TX'


source('NovTX.gffcompare.R')
```

```{r bib, include=FALSE}
# KEEP THIS AT THE END OF THE DOCUMENT TO GENERATE A LOCAL bib FILE FOR PKGS USED
knitr::write_bib(sub("^package:", "", grep("package", search(), value=TRUE)), file='skeleton.bib')
```